### 2020-04-28 Current Version/Untitled.tex
reorganize

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{bioarxiv} % Sets Journal name

\usepackage{lineno} % add
  \linenumbers % turns line numbering on
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A White Noise Approach to Evolutionary Ecology},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% Pandoc header
\usepackage{xr}
\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage{hyperref}
\usepackage{amsmath}
\externaldocument{SM}
\modulolinenumbers[2]
\usepackage{float}
\usepackage{flafter}

\begin{document}
\begin{frontmatter}

  \title{A White Noise Approach to Evolutionary Ecology}
    \author[Bio]{Bob Week\corref{1}}
   \ead{bobweek@gmail.com} 
    \author[Stats]{Steve Krone}

    \author[Bio]{Scott L. Nuismer}

    \author[Bio]{Luke J. Harmon}

      \address[Bio]{University of Idaho, Department of Biological Sciences, 875 Perimeter
Drive MS 3051, Moscow, Idaho, 83844}
    \address[Stats]{University of Idaho, Department of Statistical Science, 875 Perimeter
Drive MS 1104, Moscow, Idaho, 83844}
      \cortext[1]{Corresponding Author}

  \begin{abstract}
  We derive the dynamics of the distribution of a quantitative character
  and the abundance of a biological population from a stochastic partial
  differential equation driven by space-time white noise. In the process
  we develop a useful set of heuristics to operationalize the powerful,
  but abstract theory of white noise and measure-valued stochastic
  processes. This approach allows us to compute the full implications of
  demographic stochasticity on phenotypic distributions and abundances of
  populations. We demonstrate the utility of our approach by deriving a
  quantitative genetic model of diffuse coevolution mediated by
  exploitative competition for a continuum of resources. Other than trait
  and abundance distributions, this model predicts interaction networks
  parameterized by rates of interactions, competition coefficients, and
  selection gradients. We briefly investigate the relationship between
  selection gradients and competition coefficients. This illustrative
  investigation suggests absolute values of linear selection gradients are
  negatively correlated with competition coefficients when the ratio of
  interspecific trait variation to intraspecific trait variation is
  approximately less than \(1.25\) and positively correlated otherwise.
  The inverse relationship holds for quadratic selection gradients except
  with threshold diversity ratio equal to unity instead of \(1.25\).
  Hence, this approach contributes to the development of a synthetic
  theory of evolutionary ecology by formalizing first principle
  derivations of stochastic models that underlie rigorous investigations
  of the relationship between feedbacks of biological processes and the
  patterns of diversity they produce.
  \end{abstract}

 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Our goal in this manuscript is to develop a rigorous, but accessible
approach to synthesize the stochastic dynamics of abundance, mean trait
and heritable variation in biological populations for the study of
theoretical evolutionary ecology. A primary aim of theoretical
evolutionary ecology is the development of mathematical approaches to
describe the evolution of populations and their interactions with both
the biotic and abiotic environments in which they are embedded. Given
this consideration, a natural scope for such an approach centers on
quantifying the abundance dynamics of populations and the evolution of
traits mediating their interactions as functions of relevant abiotic
factors. Although taking into account abundance, phenotype and
environment provides the basis for a partial understanding of the
complex nature of biological communities, a deeper understanding must
account for the effects of dispersal and the phylogeographic history of
interacting lineages (Hickerson et al., 2010; Kraft et al., 2007;
Manceau et al., 2016; McPeek, 2017) along with the genetic basis of
ecologically relevant traits (Conner, 2004; Fussman et al., 2007) and
feedbacks between populations and the biogeochemical cycles they
ultimately depend on (Ågren and Andersson, 2012; Loreau, 2010). It is
therefore ideal that the development of any such mathematical approach
anticipates extensions to account for these important factors shaping
ecological communities, especially as empirical and conceptual work in
these directions continues to grow (Abdala-Roberts and Mooney, 2014;
Crutsinger, 2015; Fitzpatrick et al., 2015, 2017; Harmon et al., 2019;
Kölzsch et al., 2015; Marx et al., 2017; Nuland et al., 2019; Rudman et
al., 2017; Skovmand et al., 2018). Furthermore, the approach would
benefit from a stochastic component to capture the chance nature of
biological reality (Lande et al., 2003; Meester et al., 2018; Mubayi et
al., 2019) and serve as a basis for the construction of statistical
methods that measure evolutionary and ecological processes occuring in
the wild. Such methods will tether theory to reality and allow for
rigorous tests of hypotheses on the structure and behavior of ecological
communities. In this paper we introduce a framework that establishes a
formal connection between the continuous-time dynamics of abundance and
quantitative traits in stochastically evolving populations. We then
demonstrate the utility of our framework through the derivation and
analysis of a model of diffuse coevolution and discuss how it can be
extended to account for the details mentioned above.

Current theoretical approaches to synthesize evolution and ecology have
capitalized on the fact that biological fitness plays a key role in
determining both sets of dynamics. While correlation of fitness and
genotype is the basis of evolution by natural selection, the mean
fitness across all individuals in a population determines the growth,
stasis or decline of abundance. In section \ref{deterministic} we review
the mathematical formalization of this connection, which has been
established in the contexts of population genetics (Crow and Kimura,
1970; Roughgarden, 1979), evolutionary game theory (Hofbauer and
Sigmund, 1998; Lion, 2018; Nowak, 2006), quantitative genetics (Doebeli,
1996; Lande, 1982; Lion, 2018) and a unifying framework for these three
distinct approaches to evolutionary theory (Champagnat et al., 2006)
which includes our approach as a special case. We note here there are
two primary differences between the results presented in this paper and
those of Champagnat, Ferrière and Méléard (2006). First, instead of
developing a unifying approach to evolutionary theory, we focus on
developing a stochastic synthesis of population dynamics and
quantitative genetics including specific expressions for the dynamics of
abundance, mean trait and additive genetic variance. Second, although
Champagnat, Ferrière and Méléard (2006) provide rigorous proofs of
results relevant to our approach, they do not communicate the nessecary
tools we make use of for deriving the dynamics of abundance, mean trait
and additive genetic variance. We aim to communicate these tools to a
wide audience by striking a balance between intuitive appeal and
mathematical rigor.

Although reviewing previously established theoretical results reveals a
beautiful synthesis of evolution and population ecology, it also reveals
a gap in theoretical approaches to incorporate the intrinsically random
nature of populations. Specifically, in theoretical quantitative
genetics the derivation of a population's response to random genetic
drift is derived in discrete time under the assumption of constant
effective population size using arguments based on properties of random
samples (Lande, 1976). Though this approach conveniently mimics the
formalism provided by the Wright-Fisher model of population genetics,
real population sizes fluctuate over time. Furthermore, since these
fluctuations are themselves stochastic, it seems natural to derive
expressions for the evolutionary response to demographic stochasticity
and consider how the results relate to characterizations of random
genetic drift. This can be done in continuous time for population
genetic models without too much technical overhead, assuming a finite
number of alleles (Gomulkiewicz et al., 2017; Lande et al., 2009;
Parsons et al., 2010). However, for populations with a continuum of
types, such as a quantitative trait, finding a formal approach to derive
the evolutionary response to demographic stochasticity has remained a
vexing mathematical challenge. In this paper we close this gap by
combining the calculus of white noise with results on rescaled limits of
branching Brownian motion processes (BBM) and stochastic partial
differential equations (SPDE), which are stochastic analogs of
deterministic partial differential equations (PDE). Our goal has two
components: 1) Establish a novel synthetic approach to theoretical
evolutionary ecology that provides a formal connection between
demographic stochasticity and random genetic drift in the context of
quantitative traits. 2) Communicate some useful properties of space-time
white noise, BBM and SPDE to as wide of audience as possible. With this
goal in mind we will not provide a rigorous treatment of any of these
deep subjects. Instead, we introduce a set of heuristics that only
require the basic concepts of Reimann integration, partial
differentiation and some exposure to Brownian motion and stochastic
ordinary differential equations (SDE). For a concise introduction to SDE
and Brownian motion, we recommend the primer by Evans (2014). Rigorous
treatments of SPDE and rescaled limits of BBM can be found in Walsh
(1986) and Dawson (1993) respectively.

Through accomplishing these goals this work yields five main results.
Firstly, we develop a concise, but accurate approach for understanding
SPDE and provide heuristics to empower mathematical evolutionary
ecologists to derive novel models that relax key assumptions commonly
made in combining models of evolutionary and abundance dynamics. Second,
we formally derive general SDE driven by demographic stochasticity that
express the joint dynamics of abundance, mean trait and additive genetic
variance for populations exhibiting a wide range of potential phenotypic
distributions. Although these dynamics have been considered before while
accounting for demographic stochasticity (Polechová and Barton, 2015),
to our knowledge this is the first time they have been rigorously
derived from first principles. Third, we simplify these SDE by
approximating the abundance density of phenotypes with a Gaussian curve
and integrate the results with a model of imperfect inheritance to find
formula that generalize classical quantitative genetic expressions. This
includes an expression for the response of additive genetic variance to
selection in terms of fitness gradients analogous to the expression for
the response in mean phenotype found by Lande (1976). Fourth, following
classical niche theory we provide a rigorous mechanistic derivation of
fitness for populations experiencing resource competition and abiotic
stabilizing selection using diffusion limits of measure-valued branching
processes. This calculation provides a demonstration for performing such
diffusion limits that can be applied in other contexts. Fifth, we
explore implications of the resulting model of diffuse coevolution by
investigating the relationship between selection gradients and
competition coefficients, discovering the conditions under which these
quantities positively or negatively covary.

We now summarize the basic structure of the paper. To provide motivation
for the stochastic equations developed later and background for our
model of coevolution, we begin with \S\ref{deterministic} by briefly
summarizing derivations of deterministic evolutionary and demographic
dynamics. Starting with a PDE (which we refer to as the Deterministic
Asexual Gaussian allelic model with Abundance dynamics, abbreviated
DAGA), we arrive at a general set of ordinary differential equations
modelling the dynamics of abundance, trait mean and trait variance. We
discuss the relationship between DAGA and previous models of phenotypic
evolution that have been employed in theoretical quantitative genetics.
In \S\ref{wnc} we discuss how a diffusion limit of a spatially
structured branching process (i.e., a measure-valued branching process)
leads to a SPDE that provides a natural stochastic generalization of
DAGA. We refer to this SPDE as the Stochastic Asexual Gaussian allelic
model with Abundance dynamics (abbreviated SAGA). The diffusion limit in
turn provides a rigorous method for constructing fitness functions used
in models of evolutionary ecology. However, before we discuss
measure-valued braching processes and their diffusion limits, we find it
necessary to review some fundamental results on the calculus of white
noise and introduce a set of mathematical tools based on these results.
In SM \S\ref{SDE_DERIV} we employ these tools to derive a system of SDE
generalizing our deterministic results to account for demographic
stochasticity. We find the effect of demographic stochasticity on the
evolution of mean trait and phenotypic variance characterizes the
process of random genetic drift in continuous time. However, although
biologically insightful, these equations remain difficult to analyze and
implement numerically. In \S\ref{equations} we use an assumption of
normally distributed trait values to simplify these expressions into
tractable formula that express the evolutionary response to selection in
terms of fitness gradients. We then account for the constraint of
adaptive evolution on the amount of heritable variation available in a
population by extending these results via a model of imperfect
inheritance. The resulting equations coincide with classical results in
quantitative genetics as a special case. In \S\ref{coev} we combine the
derived equations of population dynamics with classical niche theory to
formulate a model of coevolution across a guild of \(S\) species
participating in exploitative competition along a common resource
continuum. In SM \S\ref{diffuse} we apply a result on rescaled limits of
BBM to provide a rigorous derivation. In \S\ref{dynamics} we illustrate
our model of coevolution through numerical integration for \(S=100\)
species, tracking the dynamics of traits and abundances under scenarios
of weak and strong competition. We include an account of the natural
history of the simulated system and discuss the significance of
demographic stochasticity for structuring ecological communities. In
\S\ref{ecoevo} we provide expressions for selection gradients and
competition coefficients implied by our model and use these expressions
to investigate the relationship between the degree of competition and
strength of coevolution. Finally, \S\ref{conclusion} concludes with a
summary of accomplishments, a few remarks on the limits of this approach
and future directions to incorporate more explicitly the effects of
small populations, the genetic architecture of quantitative traits,
feedbacks with ecosystem processes and the macroevolutionary history of
interacting lineages.

\hypertarget{the-framework}{%
\section{The framework}\label{the-framework}}

At the core of our approach is a stochastic analog of the replicator
equation with mutation in continuous time and phenotypic space (Schuster
and Sigmund, 1983; Taylor and Jonker, 1978). From this stochastic
replicator-mutator equation we derive a system of SDE for the dynamics
of abundance, mean trait and additive genetic variance of a population.
Hence, our approach develops a quantitative genetic theory of
evolutionary ecology. A popular alternative to quantitative genetics is
the theory of adaptive dynamics. As demonstrated by Page and Nowak
(2002) and Champagnat, Ferrière and Méléard (2006), the canonical
equation of adaptive dynamics can be derived from the replicator-mutator
equation. Thus, one could start from the atomic roots of our approach
and pursue a stochastic adaptive dynamic theory instead. We choose the
former in anticipation of an extension of our approach that explicitly
models the genetic details of populations.

In this section we review the derivations of the replicator-mutator
equation and trait dynamics from abundance dynamics and extend these
formula along with related results to the stochastic case. The results
established in this section provide the framework from which larger
scale ecological stuctures, such as species abundance distributions and
interaction networks, can be computed.

\hypertarget{deterministic-dynamics}{%
\subsection{\texorpdfstring{Deterministic dynamics
\label{deterministic}}{Deterministic dynamics }}\label{deterministic-dynamics}}

Our review begins by considering the dynamics of an asexually
reproducing population in a homogeneous environment. For simplicity, we
first assume individuals are haploid and carry one of \(K\) alleles each
with a different fitness expressed as growth rate before introducing a
model involving a quantitative trait. Under these assumptions, the
derivation of the evolution of allele frequencies due to natural
selection can be derived from expressions of exponential growth. This,
and a few related approaches, have been provided by Crow and Kimura
(1970, \S 5.3). Specifically, denoting \(\nu_i\) the abundance of
individuals with allele \(i\) and \(m_i\) the growth rate of allele
\(i\) (called the Malthusian parameter in Crow and Kimura, 1970), we
have

\begin{equation}\label{first}
\frac{d\nu_i}{dt}=m_i\nu_i.
\end{equation}

Starting from this model, we get the total abundance of the population
as \(N=\sum_{i=1}^K\nu_i\), the frequency of allele \(i\) as
\(p_i=\nu_i/N\) and the mean fitness of the population as
\(\bar m=\sum_{i=1}^Kp_im_i\). Hence, we can employ some elementary
calculus to derive the dynamics of abundance \(dN/dt\) and the dynamics
of allele frequencies \(dp_1/dt,\dots,dp_K/dt\) as

\begin{equation}
\frac{dN}{dt}=\sum_{i=1}^K\nu_im_i=N\sum_{i=1}^Kp_im_i=\bar mN,
\end{equation} \begin{equation}\label{popgen}
\frac{dp_i}{dt}=\frac{d}{dt}\frac{\nu_i}{N}=\frac{1}{N^2}\left(N\frac{d\nu_i}{dt}-\frac{dN}{dt}\nu_i\right)
=\frac{1}{N}\left(m_i\nu_i-\bar m Np_i\right)=(m_i-\bar m)p_i.
\end{equation}

Two important observations of these equations include (i) mean fitness
\(\bar m\) determines the abundance dynamics of the entire population
and (ii) allele \(i\) will increase (decrease) in frequency if
\(m_i>\bar m\) (\(<\bar m\)). Equation (\ref{popgen}) is known in the
field of evolutionary game theory as the replicator equation (Hofbauer
and Sigmund, 1998; Lion, 2018; Nowak, 2006; Schuster and Sigmund, 1983;
Taylor and Jonker, 1978). Instead of being explicitly focused on
alleles, the replicator equation describes the fluctuations of relative
abundances of various \emph{types} in a population in terms of the vital
rates of each type. Using a matrix of transition rates between differing
types, it is straight-forward to extend the replicator equation to
include mutation, which is known as the replicator-mutator equation
(Nowak, 2006).

Inspired by equations (\ref{first})-(\ref{popgen}), we derive an analog
of the replicator-mutator equation for a continuum of types (that is,
for a quantitative trait). In particular, we model a continuously
reproducing population with trait values \(x\in\mathbb{R}\) and an
abundance density \(\nu(x,t)\) that represents the amount of individuals
in the population with trait value \(x\) at time \(t\). Hence, the
abundance density satisfies \(N(t)=\int_{-\infty}^{+\infty}\nu(x,t)dx\)
and \(p(x,t)=\nu(x,t)/N(t)\) is the relative density of trait \(x\)
which we also refer to as the phenotypic distribution.

To stay within the realm of biological plausibility we require a set
technical assumptions. First, we assume the initial abundance density is
continuous, non-negative, integrable and has finite trait mean and
variance. That is, we assume \(\nu(x,0)\) is continuous in \(x\),
satisfies \(\nu(x,0)\geq0\) for all \(x\in\mathbb{R}\) and

\begin{equation}
N(0)=\int_{-\infty}^{+\infty}\nu(x,0)dx<+\infty,
\end{equation} \begin{equation}
-\infty<\bar x(0)=\int_{-\infty}^{+\infty}xp(x,0)dx<+\infty,
\end{equation} \begin{equation}
\sigma^2(0)=\int_{-\infty}^{+\infty}(x-\bar x(0))^2p(x,0)dx<+\infty,
\end{equation}

where \(\bar x(t)\) and \(\sigma^2(t)\) are respectively the mean trait
and phenotypic variance at time \(t\geq0\). Second, we assume selection
is determined by the growth rate \(m(h,x)\) that is differentiable with
respect to both arguments and is bounded above by some
\(r\in\mathbb{R}\) for all \(x\in\mathbb{R}\) and \(h\geq0\). We also
assume \((1+h)C\leq m(h,x)\) for some constant \(C>0\) so that
\(m(h,x)\) is linear or more than linear in \(h\). We denote by \(K\) an
operator that takes functions such as \(\nu(x,t)\) as its argument and
returns other functions that may still depend on the spatial variable
\(x\) and the temporal variable \(t\). For our application we will be
focused on operators \(K\) that can be written as
\((K\nu)(x,t)=\int_{-\infty}^{+\infty}\kappa(x-y)\nu(y,t)dy\) for some
non-negative and bounded function \(\kappa\). We set \(h=(K\nu)(x,t)\)
to account for nonlocal effects of the abundance density \(\nu(x,t)\) on
the fitness of individuals with trait value \(x\) (Volpert, 2014). As an
example of a nonlocal effect, consider a population experiencing
intraspecific competition for a common resource such that the
competitive interactions are not mediated by any set of traits. In this
case all individuals in the population effect the fitness of one another
even though they may exhibit a diverse range of trait values. The term
nonlocal comes from interpreting the trait value as spatial location.
From here on we abbreviate \(m(h,x)=m\big((K\nu)(x,t),x\big)\) to just
\(m(\nu,x)\). Third, we assume mutation is captured by diffusion with
coefficient \(\frac{\mu}{2}\). With these technicalities addressed we
model the demographic dynamics of a population and the dynamics of a
quantitative character simultaneously by the PDE

\begin{equation}\label{eq1}
\frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)
\end{equation}

with the initial condition \(\nu(x,0)\) described above. Equation
(\ref{eq1}) is a semilinear parabolic PDE and is referred to as a scalar
reaction-diffusion equation (Evans, 2010; Zheng, 2004). To ensure
existence and uniqueness of solutions to (\ref{eq1}), we further assume
that for each \(M>0\) there exists a constant \(L_M>0\) depending on
\(M\) such that
\(\int_\mathbb{R}\nu_1(x)dx,\int_\mathbb{R}\nu_2(x)dx\leq M\) implies

\begin{equation}\label{local_lipschitz}
\int_\mathbb{R}\left|m(\nu_1,x)\nu_1(x)-m(\nu_2,x)\nu_2(x)\right|dx\leq L_M\int_\mathbb{R}|\nu_1(x)-\nu_2(x)|dx.
\end{equation}

In SM \S\ref{finite} we combine the above assumptions on growth rate,
mutation and initial condition to prove that solutions to (\ref{eq1})
satisfy \(N(t),|\bar x(t)|,\sigma^2(t)<+\infty\) for all \(t\geq0\).

Interpreting the trait value \(x\) as spatial location and ignoring
nonlocal effects, which can be satisfied by setting \(K\nu=\nu\), this
model of spatially distributed population dynamics has been intensely
studied (for a review, see Cantrell and Cosner, 2004). When \(\mu=0\),
equation (\ref{eq1}) can be seen as an analog of equation (\ref{first})
for a continuum of types. By assuming mutation acts as diffusion, the
effect of mutation causes \(\nu(x,t)\) to flatten out over time. In
fact, if the growth rate is constant across \(x\), then this model of
mutation will cause \(\nu(x,t)\) to converge to a flat line as
\(t\to\infty\).

Although clearly an idealized representation of biological reality, this
model is sufficiently general to capture a large class of dynamics
including density dependent growth and frequency dependent selection. As
an example, logistic growth combined with quadratic stabilizing
selection can be captured using the growth rate

\begin{equation}\label{special_m}
m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_{-\infty}^{+\infty}\nu(y,t)dy
\end{equation}

where \(r\in\mathbb{R}\) is the intrinsic rate of growth in the absence
of abiotic stabilizing selection, \(a>0\) is the strength of abiotic
stabilizing selection around the phenotypic optimum
\(\theta\in\mathbb{R}\), \(c>0\) captures the sensitivity of fitness to
intraspecific competition and \(\kappa(x-y)=1\) so that competitive
interactions cause the same reduction in fitness regardless of trait
value. This exemplary fitness function has a few convenient properties.
First, the effect of competition induces a local carrying capacity on
the population, leading to a finite equilibrium abundance over bounded
subsets of \(\mathbb{R}\). Second, abiotic selection prevents the
abundance density from diffusing too far from the abiotic optimum,
leading to a finite equilibrial abundance across all of \(\mathbb{R}\),
\(\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})\),
finite equilibrial mean trait \(\lim_{t\to\infty}\bar x(t)=\theta\) and
finite equilibrial phenotypic variance
\(\lim_{t\to\infty}\sigma^2(t)=\sqrt{\frac{\mu}{a}}\), so long as
\(\bar x(0)\in\mathbb{R}\), \(\sigma^2(0)\in[0,+\infty)\),
\(N(0)\in(0,+\infty)\) and \(r>\frac{1}{2}\sqrt{a\mu}>0\). We prove
these statements in SM \S\ref{equilib}. Note in particular the
coincidence of equilibrial phenotypic variance predicted by this model
and that predicted by the classic Gaussian allelic model of quantitative
genetics (Bürger, 2000; Johnson and Barton, 2005; Lande, 1975). In
\S\ref{coev} and SM \S\ref{diffuse} we find this fitness function is
closely related to a fitness function derived from niche theory that we
use to model coevolution driven by resource competition.

To derive a replicator-mutator equation from equation (\ref{eq1}), we
employ the chain rule from calculus. Writing
\(\bar m(t)=\int_{-\infty}^{+\infty}m(\nu,x)p(x,t)dx\) for the mean
fitness, we have

\begin{multline}\label{N_det}
\frac{d}{dt}N(t)=\frac{d}{dt}\int_{-\infty}^{+\infty}\nu(x,t)dx=\int_{-\infty}^{+\infty}\frac{\partial}{\partial t}\nu(x,t)dx \\
=\int_{-\infty}^{+\infty} m(\nu,x)\nu(x,t)dx+\int_{-\infty}^{+\infty}\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)dx \\
=N(t)\int_{-\infty}^{+\infty} m(\nu,x)p(x,t)dx=\bar m(t) N(t).
\end{multline}

Using our assumptions on mutation and rate of growth, we show in SM
\S\ref{finite} \(\nu(x,t)\) is twice differentiable with respect to
\(x\) and \(\int_{-\infty}^{+\infty}\nu(x,t)dx<\infty\) for all
\(t\geq0\). This implies that we are justified in swapping the order of
differentiation and integration and the result
\(\int_{-\infty}^{+\infty}\frac{\partial^2}{\partial x^2}\nu(x,t)dx=0\)
can be derived from the fundamental theorem of calculus. Biological
reasoning agrees with this latter result since mutation neither creates
nor destroys individuals, but merely changes their type from their
parental type. Taking the same approach, we derive the dynamics of the
phenotypic distribution \(p(x,t)\) in response to selection and muation
as

\begin{multline}\label{p_PDE}
\frac{\partial}{\partial t}p(x,t)=\frac{\partial}{\partial t}\frac{\nu(x,t)}{N(t)}=\frac{1}{N^2(t)}\left(N(t)\frac{\partial}{\partial t}\nu(x,t)-\nu(x,t)\frac{d}{dt}N(t)\right) \\
=\frac{1}{N(t)}\Big(m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)-\bar m(t)\nu(x,t)\Big) \\
=\Big(m(\nu,x)-\bar m(t)\Big) p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t).
\end{multline}

This result closely resembles Kimura's continuum-of-alleles model
(Bürger, 1986; Kimura, 1965). The primary difference being that our
model utilizes diffusion instead of convolution with an arbitrary
mutation kernel. Of course, our model of mutation can be derived as an
approximation to Kimura's model, which, for sexually reproducing
populations, has been referred to as the Gaussian allelic approximation
in reference to the distribution of mutational effects at loci in a
genome on the values of traits (Lande, 1975; but see also Bürger, 2000;
along with Johnson and Barton, 2005), the infinitesimal model in
reference to modelling continuous traits as being encoded by an infinite
number of loci each having infinitesimal effect (Barton et al., 2017)
and the Gaussian descendants approximation in reference to offspring
trait values being normally distributed around their parental values
(Turelli, 2017). Alternatively, since diffusion is the continuous-time
equivalent to convolution against a Gaussian kernel (SM
\S\ref{diffconvequiv}), equation (\ref{p_PDE}) can also be seen as a
special case of Kimura's continuum-of-alleles model. We return to these
concepts in \S\ref{inheritance} where we review a simple model of
inheritance that leads us to an expression for the stochastic evolution
of additive genetic variance. To distinguish this model from previous
models of phenotypic evolution we refer to PDE (\ref{eq1}) from which
(\ref{p_PDE}) was derived as the Deterministic Asexual Gaussian allelic
model with Abundance dynamics (abbreviated DAGA). Later, we will extend
this model to include the effects of demographic stochasticity, which we
refer to as the Stochastic Asexual Gaussian allelic model with Abundance
dynamics (abbreviated SAGA).

The covariance of fitness and phenotype across the population is defined
as

\begin{equation}
\mathrm{Cov}_t\Big(m(\nu,x),x\Big)=\int_{-\infty}^{+\infty}\Big(m(\nu,x)-\bar m(t)\Big)\Big(x-\bar x(t)\Big)p(x,t)dx.
\end{equation}

Hence, the dynamics of the mean trait \(\bar x(t)\) can be derived as

\begin{multline}\label{detmean}
\frac{d}{dt}\bar x(t)=\frac{d}{dt}\int_{-\infty}^{+\infty} x p(x,t)dx=\int_{-\infty}^{+\infty} x\frac{\partial}{\partial t} p(x,t)dx \\
=\int_{-\infty}^{+\infty} x\big(m(\nu,x)-\bar m(t)\big) p(x,t)+x\frac{\mu}{2}\frac{\partial^2}{\partial x^2} p(x,t)dx
=\mathrm{Cov}_t\Big(m(\nu,x),x\Big).
\end{multline}

Equation (\ref{detmean}) is a continuous time analog of the well known
Robertson-Price equation without transmission bias (Frank, 2012; Lion,
2018; Price, 1970; Queller, 2017; Robertson, 1966). Whether or not the
covariance of fitness and phenotype creates change in \(\bar x\) to
maximize mean fitness \(\bar m\) depends on the degree to which
selection is frequency dependent (Lande, 1976). Since this change is
driven by a covariance with respect to phenotypic diversity, the
response in mean trait to selection is mediated by the phenotypic
variance. In particular, when \(\sigma^2=0\), \(\bar x\) will not
respond to selection. The result
\(\int_{-\infty}^{+\infty} x\frac{\partial^2}{\partial x^2} p(x,t)dx=0\)
can be found by applying integration by parts. Following the approach
taken to calculate the evolution of \(\bar x\), we find the response of
phenotypic variation to this model of selection and mutation is

\begin{equation}\label{vardyn_det}
\frac{d}{dt}\sigma^2(t)=\mathrm{Cov}_t\Big(m(\nu,x),(x-\bar x)^2\Big)+\mu.
\end{equation}

For the sake of space we relegate the derivation of \(d\sigma^2/dt\) to
SM \S\ref{var_deriv}. In the absence of mutation equation
(\ref{vardyn_det}) agrees with the result derived by Lion (2018) for
discrete phenotypes. From a statistical perspective, if we think of
\((x-\bar x)^2\) as a square error, then in analogy to the dynamics of
the mean trait, we see that the response in \(\sigma^2\) to selection
can be expressed as a covariance of fitness and square error, which is
defined in analogy to \(\mathrm{Cov}_t(m(\nu,x),x)\). Just as for the
evolution of \(\bar x(t)\), this covariance also creates change in
\(\sigma^2\) that can either increase or decrease mean fitness
\(\bar m\), depending on whether or not individual fitness depends on
the distribution of traits across the entire population. The effect of
selection on phenotypic variance can be positive or negative depending
on whether selection is stabilizing or disruptive.

In SM \S\ref{SDE_DERIV}, we extend these results to include the effects
of demographic stochasticity. The key realization in this direction is
that a rigorous derivation of a SPDE that naturally generalizes DAGA and
accounts for ecological interactions has been provided by Li (1998) and
slightly generalized by Champagnat, Ferriére and Méléard (2006). To
understand how this SPDE was derived, we provide an informal discussion
of measure-valued branching processes (which treat populations as sets
of discrete individuals) and their diffusion limits, the so-called
superprocesses (which treat population size as a continuously varying
number). Similar to the approach taken by Champagnat, Ferriére and
Méléard (2006), we begin with a BBM that models populations as discrete
sets of reproducing individuals whose vital rates depend on their trait
value as well as the state of the entire population. Taking a large
population size limit and keeping our assumption of single dimensional
traits, we employ a pair of classical results that show, under the
appropriate rescaling in time, phenotypic space and population density,
a sequence of rescaled BBM converges to a limiting process that can be
characterized by a SPDE (Champagnat et al., 2006; Li, 1998; Méléard and
Roelly, 1993, 1992). The limiting processes of rescaled BBM have been
referred to as measure-valued Markov processes (Dawson, 1993) or
superprocesses (Etheridge, 2000). Under the simplifying assumptions
inherited from our treatment of deterministic dynamics and the
additional assumption that the magnitude of demographic stochasticity is
independent of trait values, we obtain as a special case a relatively
simple expression for an SPDE that generalizes DAGA. The simplicity of
our special case allows us to use properties of space-time white noise
processes to derive a set of SDE that generalize equations
(\ref{N_det}), (\ref{detmean}) and (\ref{vardyn_det}) to include the
effects of demographic stochasticity. Classical expressions for the
effects of random genetic drift on the evolution of mean traits are
obtained as a further special case.

In the following section we provide the necessary mathematical tools
needed to derive SDE from SPDE. Since our aim is to present this
material to as wide of audience as possible, our approach deviates from
standard definitions to remove the need for a detailed technical
treatment. In addition to the notions of Reimann integration and partial
differentiation already employed, the reader will only need some
elementary probability and an intuitive understanding of SDE, including
Brownian motion. Because space-time white noise, denoted by
\(\dot W(x,t)\), appears in the SPDE characterizing diffusion limits of
BBM, we begin by defining \(\dot W(x,t)\) and illustrating its relevant
properties including a set of heuristics for performing calculations.
Treating only the simplist of cases, we then provide a brief review of
BBM, their diffusion limits and the SPDE that characterize them. For
those not interested in the white noise calculus or superprocesses and
would rather jump straight into more biologically relevant results, we
recommend skipping to \S\ref{equations}.

\hypertarget{white-noise-calculus-and-superprocesses}{%
\subsection{\texorpdfstring{White noise calculus and superprocesses
\label{wnc}}{White noise calculus and superprocesses }}\label{white-noise-calculus-and-superprocesses}}

One can think of white noise as the static seen on old television sets
or infinitely detailed random dust spread across both time and space.
From a more mathematical, yet still informal perspective, white noise
can be thought of as a stochastic process. That is, we can picture white
noise as a collection of random variables indexed by time and possibly
space. In relation to Brownian motion, denoted by \(W\), white noise can
be interpreted of as the derivative of Brownian motion with respect to
time, denoted \(\dot W\). Since Brownian motion can be thought to take
infinitesimally small Gaussian distributed jumps at each time point,
this leads to the conceptualization of white noise as a collection of
independent Gaussian distributed random variables. Figure \ref{wn}
illustrates realizations of this conceptualized white noise in one
(left) and two (right) dimensions.

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{wn}

}

\caption{\label{wn}Approximations of sample paths of temporal white noise (left) and space-time white noise (right) with brightness scaled to value.}\label{fig:unnamed-chunk-2}
\end{figure}

However, it turns out that realizations of white noise do not exist as
functions in the classical sense. Indeed, since Brownian motion is
nowhere differentiable with respect to time, white noise cannot be
formally understood as a time derivative. Thus our notation \(\dot W\)
is only meant to aid intuition and not be taken as formal. A formal
understanding is possible by considering white noise as a
\emph{measure}-valued process (Dawson, 1975; Walsh, 1986) or as a
\emph{generalized} process that acts on classically defined functions or
stochastic processes to return either random variables or stochastic
processes (Da Prato and Zabczyk, 2014; Krylov and Rozovskii, 1981).
Since a measure-valued process can be defined from a generalized process
and a generalized process can be defined from a measure-valued process,
the distinction between the two is more or less a matter of perspective.
However, we find the perspective of white-noise as a generalized process
to be a more efficient route for developing heuristics to help with some
routine calculations involved with deriving SDE from SPDE. Hence, the
notion of a generalized process provides the general idea implemented
here. Although the treatments of SPDE provided by Krylov and Rozovskii
(1981) and Da Prato and Zabczyk (2014) extend the theory of SDE to
formally treat SPDE in a general and elegant fashion, they require the
navigation of an enourmous amount of technical definitions and detailed
proofs. To extract some particularly useful results from this theory
relevant to our goal of synthesizing the stochastic dynamics of
biological populations, we provide a streamlined approach by
capitolizing on the solid ground these authors have established. For
instance, instead of rigorously proving properties of white noise, we
simply define them to be so, taking solice in the fact that the
technical details have been worked out elsewhere. In SM \S\ref{DPZ14} we
show how our informal treatment is related to the rigorous treatment
provided by Da Prato and Zabczyk (2014).

Before diving in, we shed a bit of light on the idea of a generalized
process. A generalized process is the stochastic analog of a generalized
function, such as the Dirac delta function \(\delta\). Often one sees
\(\delta\) defined as a function satisfying the properties
\(\delta(x)=0\) for \(x\neq0\) and
\(\int_{-\infty}^{+\infty}\delta(x)dx=1\). However, since there is no
function that satisfies these properties, we refer to \(\delta\) as a
generalized function. An alternative definition of \(\delta\) considers
its action on classically defined functions \(f\). In particular,
\(\delta(f)=f(0)\), which can be heuristically written as
\(\int_{-\infty}^{+\infty}f(x)\delta(x)dx=f(0)\). Similarly, other
generalized functions can be defined by their action on classically
defined functions. Then, just as a generalized function operates on
classical functions to return a value, a generalized process acts on a
set of functions (or processes) to return a random variable (or a
classically defined stochastic process). For a brief primer on the
theory of generalized functions, see the addendum to chapter 3 of
Kolmogorov and Fomin (1999).

\hypertarget{definition-and-basic-properties-of-white-noise}{%
\subsubsection{\texorpdfstring{Definition and basic properties of white
noise
\label{wnc_intro}}{Definition and basic properties of white noise }}\label{definition-and-basic-properties-of-white-noise}}

Throughout this section, we minimize notation by writing
\(\int_{\mathbb{R}}f(x)dx=\int_{-\infty}^{+\infty}f(x)dx\) and similarly
\(\int_Df(x)dx\) for the integral of \(f\) over \(D\subset\mathbb{R}\).
We define \(\mathscr{N}_2\) as the set of stochastic processes
\(f(x,t)\) that are continuous in \(t\) and satisfy
\(\mathbb{E}\left(\int_0^t\int_\mathbb{R}f^2(x,s)dxds\right)<+\infty\)
for each \(t\geq0\). The operator \(\mathbb{E}\) denotes expectation
with respect to the underlying probability space. For each \(t\geq0\) we
set

\begin{equation}
\|f\|_t=\sqrt{\mathbb{E}\left(\int_0^t\int_\mathbb{R}f^2(x,s)dxds\right)},
\end{equation}

and make use of the convention \(f=g\) if \(\|f-g\|_t=0\) for all
\(t\geq0\). Later, when we investigate stochastic dynamics of the
abundance density \(\nu(x,t)\), we will find
\(\sqrt\nu\in\mathscr{N}_2\). This enables us to utilize the heuristics
developed in this section for the derivation of SDE describing the
stochastic dynamics of \(N(t)\), \(\bar x(t)\) and \(\sigma^2(t)\).

We define a generalized stochastic process \(\mathbf W\) that maps
processes \(f\in\mathscr{N}_2\) to real-valued stochastic processes
indexed by time \(t\geq0\), but not by space. To evaluate \(\mathbf W\)
for a process \(f\in\mathscr{N}_2\) and some time \(t\geq0\) we write
\(\mathbf W_t(f)\). Specifically, for any \(f,g\in\mathscr{N}_2\), we
define \(\mathbf W(f)\) and \(\mathbf W(g)\) to be Gaussian processes
satisfying, for any \(t,t_1,t_2\geq0\),

\begin{equation}\label{exp_WN}
\mathbb{E}\big(\mathbf W_t(f)\big)=\mathbb{E}\big(\mathbf W_t(g)\big)=0,
\end{equation} \begin{equation}\label{cov_WN}
\mathbb{C}\big(\mathbf W_{t_1}(f),\mathbf W_{t_2}(g)\big)=\mathbb{E}\left(\int_0^{t_1\wedge t_2}\int_\mathbb{R} f(x,s)g(x,s)dxds\right),
\end{equation}

where \(t_1\wedge t_2=\min(t_1,t_2)\) and \(\mathbb{C}\) denotes
covariance with respect to the underlying probability space. In
particular, denoting \(\mathbb{V}\) the variance operator with respect
to the underlying probability space, we have
\(\mathbb{V}\big(\mathbf W_t(f)\big)=\|f\|_t^2\) for all \(t\geq0\) and
\(f\in\mathscr{N}_2\). The operators \(\mathbb{E}\) and \(\mathbb{C}\)
are to be distinguished from \(\bar f(t)\) and \(\mathrm{Cov}_t(f,g)\)
which denote expectation and covariance with respect to phenotypic
diversity at time \(t\geq0\).

Since Gaussian processes are characterized by their expectations and
covariances and since we assume the \(\mathscr{N}_2\) processes are
continuous in time, the processes \(\mathbf W(f)\) and \(\mathbf W(g)\)
are well defined. As an example, if \(f\in\mathscr{N}_2\) is independent
of time, then \(\mathbf W(f)\) is a Brownian motion with variance at
time \(t\geq0\) equal to
\(\|f\|_t^2=t \ \mathbb{E}(\int_\mathbb{R}f^2(x,0)dx)\). With the
generalized process \(\mathbf W\) defined, we define the space-time
white noise \(\dot W(x,t)\) implicitly via the stochastic integral

\begin{equation}\label{informal}
``\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds"=\mathbf W_t(f), \ \forall \ f\in\mathscr{N}_2, \ t\geq0.
\end{equation}

We place quotations in the above expression to emphasize its informal
nature and that it should not be confused with classical Riemann
integration. Following this definition of white noise, we compute its
value by sampling it using \(\mathscr{N}_2\) processes. For example,
integrating white noise over a region \(D\times[0,t]\), with \(t>0\) and
\(D\) a bounded subset of \(\mathbb{R}\), is equivalent to evaluating
\(\mathbf W_t(I_D)\) for the deterministic process

\begin{equation}
I_D(x,t)=\left\{\begin{matrix}
0, & x\notin D \\
1, & x\in D
\end{matrix}\right..
\end{equation}

Since

\begin{equation}
\|I_D\|_t^2=\mathbb{E}\left(\int_0^t\int_\mathbb{R}I_D^2(x,s)dxds\right)=t\int_Ddx=t|D|<+\infty, 
\end{equation}

where \(|D|\) denotes the length of \(D\), we have
\(I_D\in\mathscr{N}_2\). Thus, using equations (\ref{exp_WN}) and
(\ref{cov_WN}) and adopting the informal notation introduced in equation
(\ref{informal}), we can write the following

\begin{equation}
\mathbb{E}\left(\int_0^t\int_D\dot W(x,s)dxds\right)=0,
\end{equation} \begin{equation}
\mathbb{V}\left(\int_0^t\int_D\dot W(x,s)dxds\right)=t|D|.
\end{equation}

Using this informal notation, equations (\ref{exp_WN}) and
(\ref{cov_WN}) can be rewritten as

\begin{equation}\label{exp_WN_xi}
\mathbb{E}\left(\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds\right)=0,
\end{equation} \begin{multline}\label{cov_WN_xi}
\mathbb{C}\left(\int_0^{t_1}\int_\mathbb{R}f(x,s)\dot W(x,s)dxds,\int_0^{t_2}\int_\mathbb{R}g(x,s)\dot W(x,s)dxds\right) \\
=\int_0^{t_1\wedge t_2}\int_\mathbb{R}f(x,s)g(x,s)dxds.
\end{multline}

To relate these formula to the common notation used for SDE, we write

\begin{equation}
\hat f(x,t)=\frac{f(x,t)}{\sqrt{\int_\mathbb{R}f^2(y,t)dy}} \ \text{ and } \ 
d\hat{\mathbf W}_t(f)=\left(\int_\mathbb{R}\hat f(x,t)\dot W(x,t)dx\right)dt
\end{equation}

so that

\begin{equation}
\int_0^td\hat{\mathbf W}_s(f)=\int_0^t\int_\mathbb{R}\frac{f(x,s)}{\sqrt{\int_\mathbb{R}f^2(s,y)dy}}\dot W(x,s)dxds.
\end{equation}

This implies

\begin{equation}
\mathbb{E}\left(\int_0^td\hat{\mathbf W}_s(f)\right)=0, \ \mathbb{C}\left(\int_0^{t_1}d\hat{\mathbf W}_s(f),\int_0^{t_2}d\hat{\mathbf W}_s(f)\right)=t_1\wedge t_2
\end{equation}

and in particular, as a function of \(t\),
\(\int_0^td\hat{\mathbf W}_s(f)\) is a standard Brownian motion for any
\(f\in\mathscr{N}_2\). Hence, \(d\hat{\mathbf W}_t(f)\) is analogous to
the traditional shorthand used to denote stochastic differentials. Thus,
equation (\ref{cov_WN_xi}) effectively extends Itô's multiplication
table to:

\begin{table}[H]
\centering\caption{An extension of It\^o's multiplication table.}\vspace{0.2cm}
\begin{tabular}{l|lll}
             & $d\hat{\mathbf W}_t(f)$               & $d\hat{\mathbf W}_t(g)$                & $dt$ \\ \hline
             &                            &                            &      \\
$d\hat{\mathbf W}_t(f)$ & $dt$                       & $\left(\int_\mathbb{R}\hat f(x,t)\hat g(x,t)dx\right)dt$ & $0$  \\
             &                            &                            &      \\
$d\hat{\mathbf W}_t(g)$ & $\left(\int_\mathbb{R}\hat f(x,t)\hat g(x,t)dx\right)dt$ & $dt$                       & $0$  \\
             &                            &                            &      \\
$dt$         & $0$                        & $0$                        & $0$
\end{tabular}\label{mult}
\end{table}

The extension of Itô's multiplication table and properties of white
noise outlined in this subsection provide a useful set of tools for
working with SPDE. In SM \S\ref{SDE_DERIV} we employ these tools to
derive SDE that track the dynamics of abundance, mean trait and
phenotypic variance of a population from a particular SPDE. In the
following subsection, we review how this particular SPDE naturally
arises as the diffusion limit of a BBM.

\hypertarget{from-branching-processes-to-spde}{%
\subsubsection{\texorpdfstring{From branching processes to SPDE
\label{stochastic}}{From branching processes to SPDE }}\label{from-branching-processes-to-spde}}

In real populations individuals are born and potentially reproduce
before they ultimately die. These three events provide the basic
ingredients of a branching process. Mathematical investigations of such
processes have a relatively deep history (Kendall, 1966). The most
simple branching process, known as the Galton-Watson process, describes
the number of individuals alive at a given time \(t\geq0\) as a
non-negative integer (Kimmel and Axelrod, 2015). Feller (1951)
introduced a formal method to approximate branching processes with
diffusion processes which are continuous in state (i.e., population size
is approximated as a continuous quantity). Since diffusion processes
possess greater analytical tractability than branching processes,
Feller's method, known as the diffusion limit, has acquired immense
popularity particularly in the field of mathematical population genetics
(Ewens, 2004). For over the past half of a century a great deal of
accomplishments have been achieved in formalizing the diffusion limits
of branching processes that describe populations of individuals occuring
in some continuous space (Barton and Etheridge, 2019; Bertoin and Le
Gall, 2003; Dawson, 1978, 1975; Etheridge, 2008; Li, 1998; Méléard and
Roelly, 1993; Perkins, 1995, 1992; Watanabe, 1968). This space can
represent geographic space or, relevant to our context, phenotypic
space. In the following subsection, we describe the BBM process, which
is a particularly important branching process with spatial structure.
This process has been very useful in the study of SPDE due to its
simplifying assumption that individuals do not interact. However, this
assumption imposes an unfortunate restriction by precluding the
modelling of ecological interactions. We therefore follow our discussion
of BBM with a review of a few important results on spatially structured
branching processes that account for interactions.

\emph{Branching Brownian motion (BBM):} A BBM tracks individuals
navigating \(d\)-dimensional Euclidean space that reproduce and senesce
between exponentially distributed intervals. Unlike other stochastic
processes that take values in \(\mathbb{R}^d\), BBM takes values in the
set of \emph{non-negative finite measures} over \(\mathbb{R}^d\).
Intuitively, one can think of a finite measure as a function that maps
subsets of \(\mathbb{R}^d\) to real numbers. To be formal, we only
consider the Borel subsets of \(\mathbb{R}^d\) corresponding to the
Euclidean metric, but understanding this technicality is not crucial to
our discussion. In particular, denoting \(X_t\) a BBM, for a subset
\(D\subset\mathbb{R}^d\), \(X_t(D)\) returns the (random) number of
individuals alive within the region \(D\) at time \(t\geq0\). The BBM
has three main components:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  \textbf{Branching rate:} In our formulation of BBM we assume Lifetimes
  of individuals are exponentially distributed with death rate
  \(\lambda>0\) and reproduction occurs simultaneously with death.
  Biologically, this implies individuals are semelparous An alternative
  formulation treats birth and death events separately to model
  iteroparity. However, under the appropriate rescaling, both approaches
  converge to the same diffusion limit (Champagnat et al., 2006). We
  therefore choose the former approach for the sake of simplicity.
\item
  \textbf{Reproductive law:} When a birth event occurs we assume a
  random (possibly zero) number of offspring are produced. The
  distribution of offspring left is called the reproductive law or
  branching mechanism. We denote the mean and variance in reproductive
  output by \(\mathscr{W}\) and \(V\) respectively. The case of
  \(\mathscr{W}=1\) is referred to as the critical condition. Since
  \(X_t(\mathbb{R}^d)\) is a Galton-Watson branching process, it is well
  known the critical condition implies extinction in finite time with
  probability one, given finite \(X_0(\mathbb{R}^d)\) (Athreya and Ney,
  1972).
\item
  \textbf{Spatial movement:} Each offspring is born at the current
  location of their parent. Immediately after birth they move around
  space according to \(d\)-dimensional Brownian motion with diffusion
  parameter \(\sqrt\mu\). In our context we interpret spatial movement
  as mutation so that the location of an individual at death represents
  the value of its phenotype. Then an individual born at location
  \(x\in\mathbb{R}^d\) that lives for \(\tau>0\) units of time will have
  a normally distributed trait centered on \(x\) with covariance matrix
  equal to \(\tau\mu\) times the \(d\times d\) identity matrix. Hence,
  offspring inherit normally distributed traits centered on their
  parental trait. This fact creates a vital link to the deterministic
  dynamics reviewed above. Indeed, in the absence of selection, the
  deterministic PDE (\ref{eq1}) reduces to the \(d=1\)-dimensional
  Kolmogorov forward equation for a Brownian motion with diffusion
  parameter \(\sqrt\mu\).
\end{enumerate}

To obtain a SPDE from a BBM we take a diffusion limit. There are several
ways to do this, but a simple approach is to rescale the mass of
individuals and time by \(1/n\), diffusion by \(\mu\to\mu/n\), branching
rate by \(\lambda\to n\lambda\), fitness by
\(\mathscr{W}\to\mathscr{W}^{1/n}\) and consider the limit as
\(n\to\infty\). Denoting the rescaled process by \(X_t^{(n)}(D)\), the
limiting process \(\mathscr{X}_t=\lim_{n\to\infty}X_t^{(n)}\) is called
a super-Brownian motion and is also a non-negative finite measure-valued
process (Watanabe, 1968). However, instead of returning the number of
individuals alive in a region of space, super-Brownian motion returns
the \emph{mass} of the population concentrated in a region of space.
Since we have rescaled individual mass by \(1/n\) and took the limit
\(n\to\infty\), individuals are no longer discrete units. Instead, the
particle view of the population gets replaced by a blob spread across
\(d\)-dimensional space. In particular, the value of
\(\mathscr{X}_t(D)\) is a continuously varying non-negative random
variable for any \(t\geq0\) and \(D\subset\mathbb{R}^d\).

Unfortunately, just as with cream cheese spread across too much toast,
the blob perspective of the population may exhibit some complicated
spatial discontinuities. However, for spatial dimension \(d=1\), it
turns out that \(\mathscr{X}_t\) is absolutely continuous with respect
to the Lebesgue measure for each \(t\geq0\) (Konno and Shiga, 1988;
Reimers, 1989). This means that we can write
\(\mathscr{X}_t(D)=\int_D \nu(x,t)dx\) for some density process
\(\nu(x,t)\). Setting \(\lambda=1\) and \(m=\ln\mathscr{W}\) this
density process satisfies the SPDE

\begin{equation}\label{neutral_SPDE}
\frac{\partial}{\partial t}\nu(x,t)=m\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)+\sqrt{V\nu(x,t)}\dot W(x,t).
\end{equation}

Since \(\nu(x,t)\) is not generally differentiable in \(x\) or \(t\),
the derivatives in expression (\ref{neutral_SPDE}) are taken in the
\emph{weak} sense (for a definition see Evans, 2010). That is, to
rigorously interpret SPDE (\ref{neutral_SPDE}), we integrate the
solution \(\nu(x,t)\) against functions \(f\in C_b^2(\mathbb{R})\) where
\(C_b^2(\mathbb{R})\) is the set of bounded and twice continuously
differentiable functions on \(\mathbb{R}\). Hence, equation
(\ref{neutral_SPDE}) is just an abbreviation for

\begin{multline}\label{neutral_int}
\int_\mathbb{R}\nu(x,t)f(x)dx-\int_\mathbb{R}\nu(x,0)f(x)dx=\int_0^t\int_\mathbb{R}\nu(x,s)\left(mf(x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}f(x)\right)dsdx \\
+\int_0^t\int_\mathbb{R}f(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds, \ \forall \ f\in C_b^2(\mathbb{R}).
\end{multline}

This expression is referred to as the \emph{mild} solution of
(\ref{neutral_SPDE}). For more on the general theory of SPDE see Walsh
(1986) and Da Prato and Zabczyk (2014). Note that since \(\nu(x,t)\) is
the density of a finite measure, it is integrable for each \(t\geq0\).
Thus, since for some \(M>0\), \(|f(x)|\leq M\) for every
\(x\in\mathbb{R}\), setting \(\varphi(x,t)=f(x)\sqrt{V\nu(x,t)}\)
implies \(\varphi\in\mathscr{N}_2\). Hence, the white noise integral on
the right-hand side of equation (\ref{neutral_int}) can be understood
using the heuristics introduced above. Evaluating equation
(\ref{neutral_int}) in the particular case of \(f(x)\equiv1\) returns
the total mass process, which we refer to as the total abundance
\(N(t)\).

A convergence theorem for the diffusion limit of a generalization of BBM
was established by Watanabe (1968). Dawson (1975) suggested that, for
spatial dimension \(d=1\), this diffusion limit should admit a density
process that satisfies a SPDE. Konna and Shiga (1988) and Reimers (1989)
independently proved Dawson's suggestion was indeed correct. The
diffusion limit of this more general branching process (in arbitrary
spatial dimension) is referred to as a Dawson-Watanabe superprocess
(Etheridge, 2000). Conditioning a Dawson-Watanabe superprocess to have
constant mass returns a Fleming-Viot process (Etheridge and March, 1991;
Perkins, 1991) which has been popular in studies of spatial population
genetics. In particular, an extension of the Fleming-Viot process, known
as the \(\Lambda\)-Fleming-Viot process, was introduced by Bertoin and
Le Gall (2003) and coined by Etheridge (2008) where it was used to
resolve some technical challenges in modelling isolation by distance
(Felsenstein, 1975; see also Barton et al., 2013; and Barton and
Etheridge, 2019). Although this provides an impressive list of
accomplishments, the Dawson-Watanabe superprocess falls short of our
needs. In particular this process assumes individuals do not interact
and thus precludes its ability to model nonlocal effects on the fitness
of individuals, such as those produced via competitive interactions.
However, this concern has been addressed, leading to constuctions of
superprocesses that account for interactions among individuals. In the
next subsection we summarize the main results in this area and introduce
the SPDE that provides the basis for our approach to theoretical
evolutionary ecology.

\emph{Interacting superprocesses:} In the above subsection we reviewed
the diffusion limit of an especially tractable measure-valued branching
process. However, we found the simplicity of this process restricts us
from modelling nonlocal effects on the fitness of individuals. Here, we
discuss superprocesses that account for such effects. The existence of
diffusion limits for a class of measure-valued branching processes
involving nonlocal effects has been treated by Méléard and Roelly (1993,
1992). The interactions can manifest as dependencies of the spatial
movement or reproductive law of individuals on their position \(x\) and
the state of the whole population described by \(X_t\). An important
result of Méléard and Roelly (1993, 1992) is a theorem that provides
sufficient conditions to construct a sequence of rescaled measure-valued
branching processes that converge to a generalization of the
Dawson-Watanabe superprocess that includes interactions. The rescaling
is analogous to that described above for non-interacting Dawson-Watanabe
superprocesses, but now the reproductive law described by
\(\mathscr{W}(X_t,x)\) and \(V(X_t,x)\), branching rate
\(\lambda(X_t,x)\) and diffusion parameter \(\sqrt{\mu(X_t,x)}\) are
allowed to depend on the whole population \(X_t\) and individual
location \(x\). In Figure \ref{rescaled} we demonstrate this rescaling
in discrete time for a population experiencing stabilizing selection and
logistic growth. Since time is discretized, the process we simulate is
formally a branching random walk. For further details on our simulation
see SM \S\ref{numerical}.

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{rescaled_plots}

}

\caption{\label{rescaled}Rescaled sample paths of a branching random walk under stabilizing selection and logistic growth. The top plot displays a sample path without scaling $(n=1)$, the middle plot shows a sample path rescaled by $n=5$ and the bottom plot shows a sample path rescaled by $n=20$.}\label{fig:unnamed-chunk-4}
\end{figure}

Interactions that manifest in the spatial movement can be used to model
mutation bias and those manisfesting in the reproductive law can model
density-dependent growth rates and frequency-dependent selection.
Perkins (1995, 1992) developed a theory of stochastic integration with
respect to the so-called \emph{Brownian trees} to characterize
interacting superprocesses and establish properties of existence and
uniqueness. Li (1998) built directly off of the construction of Méléard
and Roelly (1993, 1992) to study properties of interacting
superprocesses and, by assuming individual spatial movement occurs
independently of location \(x\) and the entire population \(X_t\),
showed the evolution of associated density processes can be described by
SPDE.

Recall \(\nu(x,t)\) denotes the density of a superprocess, given it
exists. Assuming the interactions manifest only in the reproductive law
and that spatial movement follows Brownian motion with diffusion
parameter \(\sqrt\mu\) independent of both \(X_t\) and \(x\) and the
growth rate is bounded both above and below, Li's (1998) result implies
the interacting superprocess on one dimensional trait space has a
density \(\nu(x,t)\) which is non-negative, integrable, continuous in
time and space and satisfies the SPDE

\begin{equation}\label{SPDE}
\frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)+\sqrt{V\nu(x,t)}\dot W(x,t).
\end{equation}

As mentioned above, we refer to SPDE (\ref{SPDE}) as the Stochastic
Asexual Gaussian allelic model with Abundance dynamics (abbreviated
SAGA). Comparing equation (\ref{SPDE}) to equation (3.5) of Li (1998),
our \(m\) and \(V\) correspond to Li's \(b\) and \(c\) respectively.
Using techniques developed by Evans and Perkins (1994), Champagnat,
Ferrière and Méléard (2006) generalized this result to the case when
\(m(h,x)\) is bounded above and is linear or more than linear in \(h\).
In particular, SAGA corresponds to a special case of equation (4.6) in
Champagnat, Ferrière and Méléard (2006) and is valid for the assumptions
on \(m\) made in \S\ref{deterministic}.

With solutions to SAGA well defined, we can calculate the total mass
process \(N(t)\) using the mild solution of (\ref{SPDE}) with
\(f(x)\equiv1\in C_b^2(\mathbb{R})\) (the symbol ``\(\equiv\)'' means
equal to for every \(x\)). That is,

\begin{multline}
N(t)=N(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)\cdot1+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}1\right)+1\sqrt{V\nu(x,s)}\dot W(x,s)dsdx \\
=N(0)+\int_0^t\bar m(s)N(s)dt+\int_0^t\sqrt{VN(s)}d\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)}),
\end{multline}

where

\begin{equation}
\bar m(t)=\frac{1}{N(t)}\int_\mathbb{R}m(\nu,x)\nu(x,t)dx,
\end{equation}

and

\begin{equation}
\int_0^td\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)})=\int_0^t\int_\mathbb{R}\frac{\sqrt{\nu(x,s)}}{\sqrt{\int_\mathbb{R}\nu(x,s)dx}}\dot W(x,s)dxds.
\end{equation}

Setting \(W_1(t)=\hat{\mathbf{W}}_t(\sqrt{\nu(x,t)})\), we can use
traditional stochastic differential notation to write

\begin{equation}
dN=\bar mNdt+\sqrt{VN}dW_1.
\end{equation}

To find the associated SDE for \(\bar x(t)\) and \(\sigma^2(t)\), we
want to repeat the same approach for \(f(x)=x,x^2\) and apply Itô's
lemma. However, for these cases \(f\notin C^2_b(\mathbb{R})\) since
\(f\) will not be bounded. But if we can show
\(\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty\) for all \(t>0\) given
this condition is satisfied by \(\nu(x,0)\), then we can apply the mild
solution of (\ref{SPDE}) to derive SDE for \(\bar x(t)\) and
\(\sigma^2(t)\). To illustrate, let us suppose this is the case. Setting
\(\tilde x(t)=\int_\mathbb{R}x\nu(x,t)dx\), we have

\begin{equation}
\tilde x(t)=\tilde x(0)+\int_0^t\int_\mathbb{R}\nu(x,s)m(\nu,x)x+x\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
\end{equation}

Similarly, setting
\(\tilde{\tilde\sigma}^2(t)=\int_\mathbb{R}x^2\nu(x,t)dx\), we have

\begin{equation}
\tilde{\tilde\sigma}^2(t)=\tilde{\tilde\sigma}^2(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)x^2+\mu\right)+x^2\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
\end{equation}

Since \(\bar x(t)=\tilde x(t)/N(t)\) and
\(\sigma^2(t)=\tilde{\tilde\sigma}^2(t)/N(t)-\bar x^2(t)\), we can use
Itô's lemma to derive SDE for \(\bar x(t)\) and \(\sigma^2(t)\), which
we perform in SM \S\ref{SDE_DERIV}. We make no attempt in finding
sufficient conditions to ensure
\(\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty\) and hence make no
general assertions about the existence or uniqueness of \(\bar x(t)\) or
\(\sigma^2(t)\). Regardless, we will later assume \(\nu(x,t)\) can be
approximated by a Gaussian curve in \(x\) for all \(t\geq0\). This
assumption implies \(\int_\mathbb{R}|x|^n\nu(x,t)dx<+\infty\) for all
\(n\in\mathbb{N}\) and for all \(t\geq0\).

\hypertarget{equations-of-evolutionary-and-demographic-dynamics}{%
\subsection{\texorpdfstring{Equations of evolutionary and demographic
dynamics
\label{equations}}{Equations of evolutionary and demographic dynamics }}\label{equations-of-evolutionary-and-demographic-dynamics}}

In SM \S\ref{SDE_DERIV} we show SDE for \(N(t)\), \(\bar x(t)\) and
\(\sigma^2(t)\) can be expressed as

\begin{subequations}\label{gen_eqns}
\begin{equation}\label{N}
dN(t)=\bar m(t)N(t)dt+\sqrt{V N(t)}dW_1(t),
\end{equation}
\begin{equation}\label{xbar_gen}
d\bar x(t)=\mathrm{Cov}_t\Big(x,m(\nu,x)\Big)dt+\sqrt{V\frac{\sigma^2(t)}{N(t)}}dW_2(t),
\end{equation}
\begin{equation}\label{sig2_gen}
d\sigma^2(t)=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)dt+\left(\mu-V\frac{\sigma^2(t)}{N(t)}\right)dt+\sqrt{V\frac{\overline{(x-\bar x(t))^4}-\sigma^4(t)}{N(t)}}dW_3(t),
\end{equation}
\end{subequations}

where \(W_1\), \(W_2\) and \(W_3\) are standard Brownian motions. We
note that conditions on the growth rate \(m\) to guarantee existence and
uniqueness of solutions to (\ref{xbar_gen}) and (\ref{sig2_gen}) have
yet to be investigated. However, our results on the deterministic PDE
suggest that \(m(h,x)\) bounded above and differentiable in both
arguments is sufficient. Dividing by \(dt\) one can interpret equations
(\ref{gen_eqns}) as if they are ordinary differential equations, but
this not technically rigorous since Brownian motion is nowhere
differentiable with respect to time. In SM \S\ref{SDE_DERIV} we show
that in general \(W_1\) is independent of both \(W_2\) and \(W_3\), but
\(W_2\) and \(W_3\) covary.

There is quite a bit we can learn from expressions (\ref{gen_eqns}).
Firstly, setting \(V=0\) recovers the deterministic dynamics derived in
\S\ref{deterministic}. Alternatively, one can take \(N(t)\to\infty\) to
recover the deterministic dynamics for \(\bar x(t)\) and
\(\sigma^2(t)\). Characteristically, we note the effect of demographic
stochasticity on abundance grows with \(\sqrt{N(t)}\). Hence, dividing
by \(N(t)\), we find the effects of demographic stochasticity on the
per-capita growth rate diminish with increased abundance. Relating the
response to demographic stochasticity derived here to the effect of
random genetic drift derived in classic quantitative genetic theory, if
we set \(\sigma^2(t)=\sigma^2\) and \(N(t)=N\) constant with respect to
time, then integrating the stochastic term in equation (\ref{xbar_gen})
over a single unit of time returns a normally distributed random
variable with mean zero and variance equal to \(V\sigma^2/N\). In
particular, assuming perfect inheritance, when reproductive variance is
unity (\(V=1\)) this random variable coincides with the effect of random
genetic drift on the change in mean trait over a single generation
derived using sampling arguments (Lande, 1976). There is also an
interesting connection with classical population genetics. A fundamental
result from early population genetic theory is the expected reduction in
diversity due to the chance loss of alleles in finite populations
(Fisher, 1923; Wright, 1931). This expected reduction in diversity due
to random genetic drift is captured by the third term in the
deterministic component of expression (\ref{sig2_gen}), particularly
\(-V\sigma^2(t)/N(t)\). The component of SDE (\ref{sig2_gen}) describing
random fluctuations in \(\sigma^2(t)\) is more complicated and is
proportional to the root of the difference between the centralized
fourth moment of \(p(x,t)\) and \(\sigma^4(t)\).

These expressions can be used to investigate the dynamics of the mean
and variance for general \(\nu(x,t)\), assuming
\(\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty\) for all \(t\geq0\).
However, in the next subsection we simplify these expressions by
approximating \(\nu(x,t)\) with a Gaussian curve. By assuming
\(\nu(x,t)\) is Gaussian for \(t\geq0\), we guarantee the existence of
\(\bar x(t)\) and \(\sigma^2(t)\) for all \(t\geq0\). Furthermore, in SM
\S\ref{SDE_DERIV} we show that under the Gaussian case \(W_1,W_2\) and
\(W_3\) are independent.

\hypertarget{particular-results-assuming-a-gaussian-phenotypic-distribution}{%
\subsubsection{\texorpdfstring{Particular results assuming a Gaussian
phenotypic distribution
\label{particular}}{Particular results assuming a Gaussian phenotypic distribution }}\label{particular-results-assuming-a-gaussian-phenotypic-distribution}}

By approximating \(\nu(x,t)\) with a Gaussian curve in \(x\) for each
\(t\geq0\), expressions (\ref{gen_eqns}) transform into efficient tools
for deriving the dynamics of populations given a fitness function
\(m(\nu,x)\). Gaussian phenotypic distributions are often obtained
through Gaussian, exponential or weak selection approximations together
with a simplified model of mutation and asexual reproduction or random
mating (Bürger, 2000; Lande, 1980; Turelli, 2017, 1986, 1984).
Alternatively, it has been shown that a Gaussian distribution can
provide a reasonable approximation even when selection is strong and
non-Gaussian (Turelli and Barton, 1994). Hence, given appropriate
assumptions on selection, mutation and reproduction, the abundance
density \(\nu(x,t)\) can be approximated as a Gaussian curve in \(x\)
when the ratio \(V/N\) is small (i.e., when the variance in reproductive
output is much smaller than the population size). As with any diffusion
approximation, this requires a sufficiently large abundance to
accurately reflect the dynamics of populations. Hence, models developed
in this framework are not suitable for studies involving very small
population sizes. Allowing for these restrictions, we assume

\begin{equation}
\nu(x,t)=\frac{N(t)}{\sqrt{2\pi\sigma^2(t)}}\exp\left(-\frac{\big(x-\bar x(t)\big)^2}{2\sigma^2(t)}\right).
\end{equation}

Under this assumption we find in SM \S\ref{cov2deriv} the results
(suppressing the dependency on \(t\))

\begin{equation}\label{covxm}
\mathrm{Cov}(x,m)=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right),
\end{equation} \begin{equation}
\mathrm{Cov}\Big((x-\bar x)^2,m\Big)=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)
\end{equation}

and \(\overline{(x-\bar x)^4}=3\sigma^4\). In particular, these results
imply

\begin{subequations}\label{no_inher}
\begin{equation}\label{xbar}
d\bar x=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)dt+\sqrt{V\frac{\sigma^2}{N}}dW_2,
\end{equation}
\begin{equation}\label{G}
d\sigma^2=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)dt +\left(\mu-V\frac{\sigma^2}{N}\right)dt+\sigma^2\sqrt{\frac{2V}{N}}dW_3.
\end{equation}
\end{subequations}

These equations allow us to derive the response in trait mean and
variance by taking derivatives of fitness, a much more straightforward
operation than calculating a covariance for general phenotypic
distributions. Note that in the above expressions, the partial
derivatives of \(\bar m\) represent frequency independent selection and
the averaged partial derivatives of \(m\) represent frequency dependent
selection. This relationship has already been pointed out by Lande
(1976) for the evolution of the mean trait in discrete time, but here we
see an analogous relationship holds in continuous time and also for the
evolution of trait variance.

In the next subsection we generalize this result to the case when traits
are imperfectly inherited. In this case, the phenotypic variance
\(\sigma^2\) is replaced by a genetic variance \(G\). This genetic
variance represents the component of the variance in expressed traits
\(\sigma^2\) explained by additive effects of different alleles among
genetic loci encoding for the focal phenotype (Bulmer, 1980; Lynch and
Walsh, 1998; Roughgarden, 1979). It is therefore fitting that \(G\) is
referred to as the additive genetic variance.

\hypertarget{the-evolution-of-additive-genetic-variance}{%
\subsubsection{\texorpdfstring{The evolution of additive genetic
variance
\label{inheritance}}{The evolution of additive genetic variance }}\label{the-evolution-of-additive-genetic-variance}}

To model imperfect heritability we consider the relationship between
expressed phenotypes \(x\in\mathbb{R}\) and associated genetic values
\(g\in\mathbb{R}\) known as \emph{breeding values}. The breeding value
of an individual is the sum of additive effects of the alleles carried
by the individual on its expressed trait. Since our derivations of
evolutionary equations are based on branching processes that assume
asexually reproducing populations (\S\ref{stochastic}), the additive
genetic variance \(G\) is just the variance of breeding values in a
population. For a detailed treatment of breeding values and additive
genetic variances, see Bulmer (1980) and Lynch and Walsh (1998).

Our treatment of the relationship between breeding values and expressed
traits follows classical quantitative genetic assumptions such as those
used in the seminal paper by Lande (1975) to investigate the maintenance
of genetic variation. In particular, we assume that the expressed trait
for any given individual is independent of environmental conditions and
normally distributed around their breeding value with variance \(\eta\).
Hence, \(\sigma^2=G+\eta\). In the case that all of the effects of
alleles on an expressed trait are additive, \(\eta\) is known as the
\emph{variance of environmental deviation} (Lande, 1975; Lynch and
Walsh, 1998). For a given breeding value, we denote the probability
density of a randomly drawn expressed trait by \(\psi(x,g)\). Hence,

\begin{equation}
\psi(x,g)=\frac{1}{\sqrt{2\pi\eta}}\exp\left(-\frac{(x-g)^2}{2\eta}\right).
\end{equation}

To include this relationship in our framework, we write \(\rho(g,t)\) as
the abundance density of breeding values at time \(t\) so that
\(\int_{-\infty}^{+\infty}\rho(g,t)dg=\int_{-\infty}^{+\infty}\nu(x,t)dx=N(t)\).
We switch our focus from directly modelling the evolution of
\(\nu(x,t)\) to modelling the evolution of \(\rho(g,t)\). Once
\(\rho(g,t)\) is determined, we can compute \(\nu(x,t)\) via

\begin{equation}\label{nu_from_rho}
\nu(x,t)=\int_{-\infty}^{+\infty}\rho(g,t)\psi(x,g)dg.
\end{equation}

However, since selection acts on expressed phenotypes, we use the
assumed relationship between breeding values and expressed traits to
calculate the fitness of breeding values. To motivate the approach
taken, consider the problem of inferring the breeding value of an
individual given its expressed trait \(x\). Denote \(\mathfrak{g}\) a
random variable representing the unkown breeding value. Under this model
of inheritance we know \(x\) is a random sample from a normal
distribution with mean \(\mathfrak{g}\) and variance \(\eta\).
Maximizing likelihood suggests \(x\) is our best guess for
\(\mathfrak{g}\), but the actual value of \(\mathfrak{g}\) is normally
distributed around \(x\) with the variance \(\eta\). Hence, for fixed
\(x\), we obtain \(\psi(x,g)\) as the probability density of
\(\mathfrak{g}\). Thus, the mean fitness of a breeding value \(g\)
across all individuals carrying \(g\) can be written as

\begin{equation}\label{m_rel}
m^*(\rho,g)=\int_{-\infty}^{+\infty}m(\nu,x)\psi(x,g)dx.
\end{equation}

This is similar to the approach taken by Kimura and Crow (1978) to
calculate the overall effects of selection for expressed characters onto
the changes in the distribution of alleles encoding those characters.
However, instead of focusing on the frequencies of alleles at particular
loci, our results focus on the densities of breeding values. With the
relationship between \(m(\nu,x)\) and \(m^*(\rho,g)\) established, we
define the evolution of \(\rho(g,t)\) by the SPDE

\begin{equation}\label{rho_SPDE}
\dot\rho(g,t)=m^*(\rho,g)\rho(g,t)+\frac{\mu}{2}\frac{\partial^2}{\partial^2 g}\rho(g,t)+\sqrt{V\rho(g,t)}\dot W(g,t).
\end{equation}

Equation (\ref{rho_SPDE}) is a stochastic generalization of the
deterministic PDE (\ref{eq1}) from \S\ref{deterministic}, but describes
the evolution of the distribution of breeding values instead of
expressed characters. However, whether modelling expressed characters or
breeding values, we refer to SPDE of the form (\ref{rho_SPDE}) as
Stochastic Asexual Gaussian allelic models with Abundance dynamics
(abbreviated SAGA). Replacing \(\psi(x,g)\) with different probability
densities produces different developmental models. This includes
densities that may depend on environmental variables and hence model
phenotypic plasticy. To leave room for the naming of other models where
\(\psi(x,g)\) is Gaussian, we refer to the model of development we have
chosen as the \emph{centered} Gaussian developmental model. In
\S\ref{wnc} we review the origins of equation (\ref{rho_SPDE}) and
provide some theory to help make sense of it, particularly the term
\(\dot W\).

Assuming \(\rho(g,t)\) is Gaussian implies its mode coincides with
\(\bar x\). Furthermore, since \(\sigma^2=G+\eta\), we can use equation
(\ref{m_rel}) and the chain rule from calculus (see SM \S\ref{fit2fit})
to find

\begin{subequations}
\begin{equation}
\frac{\partial\bar m}{\partial G}=\frac{\partial\bar m}{\partial\sigma^2}\frac{\partial\sigma^2}{\partial G}=\frac{\partial\bar m}{\partial\sigma^2},
\end{equation}
\begin{equation}
\overline{\frac{\partial m}{\partial G}}=\overline{\frac{\partial m}{\partial\sigma^2}\frac{\partial\sigma^2}{\partial G}}=\overline{\frac{\partial m}{\partial\sigma^2}}.
\end{equation}
\end{subequations}

Thus, equations (\ref{no_inher}) become

\begin{subequations}\label{inher}
\begin{equation}\label{xbarfinal}
d\bar x=G\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)dt+\sqrt{V\frac{G}{N}}dW_2,
\end{equation}
\begin{equation}\label{Gfinal}
dG=2G^2\left(\frac{\partial\bar m}{\partial G}-\overline{\frac{\partial m}{\partial G}}\right)dt+\left(\mu-V\frac{G}{N}\right)dt+ G\sqrt{\frac{2V}{N}}dW_3.
\end{equation}
\end{subequations}

From expressions (\ref{inher}) we see that, under this model of
inheritance, focusing on additive genetic variance \(G\) instead the
variance in expressed traits \(\sigma^2\) makes no structural changes to
the basic equations describing the dynamics of populations. In the next
section, we make use of these expressions to develop a model of diffuse
coevolution in a guild of \(S\) species competing along a resource
continuum.

\hypertarget{a-model-of-diffuse-coevolution}{%
\section{\texorpdfstring{A model of diffuse coevolution
\label{coev}}{A model of diffuse coevolution }}\label{a-model-of-diffuse-coevolution}}

\hypertarget{formulation}{%
\subsection{\texorpdfstring{Formulation
\label{form_coev}}{Formulation }}\label{formulation}}

In this section we demonstrate the use of our framework by developing a
model of diffuse coevolution across a guild of \(S\) species whose
interactions are mediated by resource competition along a single niche
axis. Because our approach treats abundance dynamics and evolutionary
dynamics simultaneously, this model allows us to investigate the
relationship between selection gradients and competition coefficients,
which we carry out in what follows.

The dynamics of phenotypic distributions and abundances have been
derived above and so the only task remaining is the formulation of a
fitness function. Our approach mirrors closely the theory developed by
MacArthur and Levins (1967), Levins (1968) and MacArthur (1972, 1970,
1969). The most significant difference, aside from allowing evolution to
occur, is the treatment of resource quality, which we replace with a
model of abiotic stabilizing selection. A derivation is provided in SM
\S\ref{diffuse}.

For species \(i\) we inherit the above notation for trait value,
distribution, average, variance, abundance, et cetera, except with an
\(i\) in the subscript. Real world examples of niche axes include the
size of seeds consumed competing finch species and the date of activity
in a season for pollinators competing for floral resources. For
mathematical convenience, we model the axis of resources by the real
line \(\mathbb{R}\). The value of a resouce along this axis is denoted
by the symbol \(\zeta\). For an individual in species \(i\), we assume
the resource utilization curve \(u_i\) can be written as

\begin{equation}
u_i(\zeta,x_i)=\frac{U_i}{\sqrt{2\pi w_i}}\exp\left(-\frac{(x_i-\zeta)^2}{2w_i}\right).
\end{equation}

We further assume the niche center \(x_i\) is normally distributed among
individuals in species \(i\), but the niche breadth \(w_i\) and total
niche utilization \(U_i\) are constant across individuals in species
\(i\) and therefore cannot evolve. Suppose \(\theta_i\in\mathbb R\) is
the optimal location along the niche axis for species \(i\) such that,
in the absence of competition, individuals leave on average \(Q_i\)
offspring when concentrated at \(\theta_i\). We assume the expected
reproductive output of an individual in species \(i\) decreases as niche
location \(\zeta\) leaves the optimum \(\theta_i\) at the rate
\(A_i>0\). Hence, abiotic stabilizing selection along the resource axis
can be modelled by the curve

\begin{equation}
e_i(\zeta)=Q_i\exp\left(-\frac{A_i}{2}(\theta_i-\zeta)^2\right).
\end{equation}

The effect of abiotic stabilizing selection on the fitness for an
individual of species \(i\) with niche location \(x_i\) is then given by

\begin{equation}
\int_{-\infty}^{+\infty}e_i(\zeta)u_i(\zeta,x_i)d\zeta=\frac{Q_iU_i}{\sqrt{A_iw_i+1}}\exp\left(-\frac{A_i}{2(A_iw_i+1)}(\theta_i-x_i)^2\right).
\end{equation}

To determine the potential for competition between individuals with
niche locations \(x_i\) and \(x_j\), belonging to species \(i\) and
\(j\) respectively, we compute the niche overlap

\begin{equation}
\mathcal O_{ij}(x_i-x_j)=\int_{-\infty}^{+\infty}u_i(\zeta,x_i)u_j(\zeta,x_j)d\zeta=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j)}}\exp\left(-\frac{(x_i-x_j)^2}{2(w_i+w_j)}\right).
\end{equation}

A notable criticism of using niche overlap to measure the intensity of
competition points to cases where populations competiting on multiple
niche axes exhibit overlap on at least one of the axes, but no overall
niche overlap (Holt, 1987). Thus niche overlap on lower-dimensional
projections of some multivariate niche space does not imply the
populations compete. To illustrate with a simple example, consider two
populations competing for space on the plane \(\mathbb{R}^2\). If the
spatial distributions of the two populations overlap, then they will
overlap on both spatial axes. However, if the populations do not overlap
on at least one of the spatial axes, they will have no overall spatial
overlap. Furthermore, even if the species overlap on both spatial axes,
they need not have any overall spatial overlap. This final result
corresponds to the fact that components of niche space do not
necessarily interact multiplicatively to determine the consequences for
the intensity of competition. In another component of Holt's (1987)
critique, an argument is made for the potential of competition occuring
without any overlap in niche space. However, this argument is based on
the practical difficulty of identifying every resource axis populations
are competing on and how these axes interact to determine fitness
consequences. Our model avoids these caveats by assuming competition
only occurs along a single dimensional resource gradient.

To map the degree of niche overlap to fitness, we assume competition
between individuals with niche locations \(x_i\) and \(x_j\) decreases
the expected reproductive output for the individual in species \(i\) at
the rate \(c_i\mathcal{O}_{ij}(x_i-x_j)\) for some \(c_i>0\). We refer
to \(c_i\) as the sensitivity to competition for species \(i\). The term
\(c_i\mathcal{O}_{ij}(x_i-x_j)\) coincides with a special case of a term
used to capture competition in Dawson's geostochastic logistic model, an
SPDE model developed to study the combined effects of demographic
stochasticity, spatial dispersion and locally finite carrying capacity
(Dawson, 1978). In relation to the example fitness function discussed in
\S\ref{deterministic}, consider
\(\kappa(x_i-x_j)=\mathcal{O}_{ij}(x_i-x_j)\). In SM \S\ref{diffuse} we
combine our treatment of resource competition with equations (\ref{N}),
(\ref{xbarfinal}) and (\ref{Gfinal}) to find

\begin{subequations}\label{comm_dynamics}
\begin{multline}\label{N_comm}
dN_i = \Bigg\{R_i-\frac{a_i}{2}\Big((\bar x_i-\theta_i)^2+G_i+\eta_i\Big) - c_i\sum_{j=1}^SN_jU_iU_j\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\Bigg\}N_idt \\ + \sqrt{V_iN_i}dW_1,
\end{multline}
\begin{multline}\label{x_comm}
d\bar x_i = \Bigg\{a_iG_i(\theta_i-\bar x_i)-c_iG_i\bigg(\sum_{j=1}^SN_jU_iU_jb_{ij}(\bar x_j-\bar x_i)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\bigg)\Bigg\}dt \\ +\sqrt{V_i\frac{G_i}{N_i}}dW_2,
\end{multline}
\begin{multline}\label{G_comm}
dG_i =  \Bigg\{c_i{G_i}^2\bigg(\sum_{j=1}^SN_jU_iU_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2} \\
+N_iU_i^2b_{ii}\sqrt\frac{b_{ii}}{2\pi}+\mu_i-a_i{G_i}^2-V_i\frac{G_i}{N_i}\bigg)\Bigg\}dt+G_i\sqrt{\frac{2V_i}{N_i}}dW_3,
\end{multline}
\end{subequations}

where

\begin{subequations}
\begin{align}
R_i = & \ \ln \left(\frac{Q_iU_i}{\sqrt{1+A_iw_i}}\right), \\
a_i = & \ \frac{A_i}{1+A_iw_i}, \\
b_{ij}(t) = b_{ji}(t) = & \ \big(w_i+w_j+\eta_i+\eta_j+G_i(t)+G_j(t)\big)^{-1}, \\
c_i \geq & \ 0.
\end{align}
\end{subequations}

Despite the convoluted appearance of system (\ref{comm_dynamics}), there
are some nice features that reflect biological reasoning. For example,
the dynamics of abundance are just a generalization of Lotka-Volterra
dynamics. In particular, the effect of competition with species \(j\) on
the fitness of species \(i\) grows linearly with \(N_j\). However, as
biotic selection pushes \(\bar x_i\) away from \(\bar x_j\), the effect
of competition with species \(j\) on the fitness of species \(i\)
rapidly diminishes, reflecting a reduction in niche overlap. The
divergence of \(\bar x_i\) and \(\bar x_j\) due to competition is
referred to in the community ecology literature as character
displacement (Brown and Wilson, 1956). We also see that the fitness of
species \(i\) drops quadratically with the difference between
\(\bar x_i\) and the abiotic optimum \(\theta_i\). Hence, abiotic
selection acts to pull \(\bar x_i\) towards \(\theta_i\). The response
in mean trait \(\bar x_i\) to natural selection is proportional to the
amount of heritable variation in the population, represented by the
additive genetic variance \(G_i\). However, we have that \(G_i\) is
itself a dynamic quantity. Under our model, abiotic stabilizing
selection erodes away heritable variation at a rate that is independent
of both \(N_i\) and \(\bar x_i\). The effect of competition on \(G_i\)
is a bit more complicated. When \(b_{ij}(\bar x_i-\bar x_j)^2<1\),
competition with species \(j\) acts as diversifying selection which
tends to increase the amount of heritable variation. However, when
\(b_{ij}(\bar x_i-\bar x_j)^2>1\), competition with species \(j\) acts
as directional selection and reduces \(G_i\). In the following
subsections we demonstrate the behavior of system (\ref{comm_dynamics})
by plotting numerical solutions and investigate implications for the
relationship between the strength of ecological interactions and
selection.

\hypertarget{community-dynamics}{%
\subsection{\texorpdfstring{Community dynamics
\label{dynamics}}{Community dynamics }}\label{community-dynamics}}

For the sake of illustration we numerically integrated system
(\ref{comm_dynamics}) for a richness of \(S=100\) species. We assumed
homogeneous model parameters across species in the community as
summarized by Table \ref{par_vals}. We repeated numerical integration
under the two scenarios of weak and strong competition. For the first
scenario of weak competition we set \(c=1.0\times10^{-7}\) and for the
second scenario of strong competition we set \(c=5.0\times10^{-6}\).
With these two sets of model parameters, we simulated our model for
\(1000.0\) units of time. For both scenarios, we initialized the trait
means to \(\bar x_i=0.0\), additive genetic variances to \(G_i=10.0\)
and abundances to \(N_i=1000.0\) for each \(i=1,\dots,S\).

Temporal dynamics for each scenario are provided in Figure
\ref{temporal}. This figure suggests weaker competition leads to
smoother dynamics and a higher degree of organization within the
community. Considering expression (\ref{N_comm}) we note that, all else
equal, relaxed competition allows for larger growth rates which promote
greater abundances. From (\ref{N_comm}) we also note that the per-capita
effects on demographic stochasticity diminish with abundance. To see
this, divide both sides by \(N_i\). Inspecting expressions
(\ref{x_comm}) and (\ref{G_comm}), we see that larger abundances also
erode the effects of demographic stochasticity on the evolution of mean
trait and additive genetic variance. These effects were already noted in
\S\ref{equations}, and thus are not a consequence of our model of
coevolution per-se, but we revisit them here since Figure \ref{temporal}
demonstrates the importance of demographic stochasticity in structuring
ecological communities even when populations are very large. Hence,
contrary to the common assumption that stochastic effects can be ignored
for large populations, we find that minute asymmetries generated by
demographic stochasticity remain significant drivers of community
structure. In particular, we initialized the species with identical
state variables and model parameters, but found an enourmous amount of
asymmetry and even some potential phase changes. In the following two
paragraphs we describe the natural history of the community as
illustrated in Figure \ref{temporal}.

We begin by describing the weak competition scenario. After about
\(125.0\) units of time, the community appears to have shaken off the
initial conditions and entered into a qualitatively distinct phase of
dynamics. Aside from a few outliers, most of the species remain
clustered together in their state variables. This lasts for
approximately \(375.0\) units of time until, at around time \(500.0\), a
drastic change occurs. At this moment the tightly packed cluster of
species begins to fan out in all three state variables. Simultaneously,
we observe large a shift in mean traits for higher values and in
additive genetic variances for lower values. Upon inspecting our
calculations, we diagnose the reason for this shift. The outlier species
that were initially pushed away from the common abiotic optimum (\(0.0\)
in this case) evolved a significant reduction in the amount of heritable
variation (\(\approx60\%\)) due to directional selection induced by
competition. This reduction in heritable variation slowed adaptation,
causing these species to linger on the outskirts of niche space, some
longer than others. In the meantime the rest of the community, being
tightly packed, experienced greater competition which led to diminished
abundances for these species and caused some members of the core group
to veer away from the abiotic optimum. The reduced abundances of the
core group led to reduced competition overall. As a result, the outlier
populations were given a slight increase in growth rate, enough to allow
them to increase their abundances orders of magnitude higher than the
species in the core group and giving them more weight in driving the
evolution of other species. Many of these heavy-hitting outlier species
had already been maintaining negative mean traits, but around time
\(500.0\) the high abundance species with positive mean traits began to
experience enough intraspecific competition to override interspecific
competition. This generated a net selection gradient and associated
evolutionary response towards the abiotic optimum. The sudden imbalance
of these high abundance species effectively induced a single large
competitive exclusion event pushing the majority of the community far
away from the abiotic optimum. After this shift the cluster began to
slowly bloom in all three state variables as species took advantage of
novel asymmetries in their competitive abilities mediated by a new
distribution of mean trait values across the community. About \(125.0\)
units of time later, the community reached a qualitatively new phase of
dynamics. If we kept running the numerical integrator, we would continue
to see similar drama unfolding over and over again as minute stochastic
changes contribute to asymmetries which slowly build into drastic
shifts.

The strong competition scenario is not quite as showy. Although the
dynamics of trait means and variances tend to be far more stochastic
than in the weak competition scenario, the community overall appears to
quickly reach some statistical equilibrium and remain there. However,
the abundances across all species in the community are very low due to
competition sensitivies being an order of magnitude higher than in the
weak competition case. Most of the species maintain abundances greater
than \(1000.0\), but we found one species that dropped to an abundance
of about \(50.0\). If we let the numerical integrator run long enough in
this case, we will likely see many of the species go extinct.

Finding ways to interpret simulated dynamics provides a useful arena to
exercise biological reasoning. However, it does not fulfill our desire
to quantify the patterns and processes present in competing communities.
In the next subsection we take a step in this direction by using our
model to derive formula for selection gradients and competition
coefficients. To investigate their relationship, we calculate their
covariances using simplifying assumptions on species abundances and
intraspecific trait variances. We then investigate how these covariances
change with the ratio of variance of interspecific mean traits to
variance of intraspecific individual traits and use a numerical approach
to investigate correlations between the strength of pairwise coevolution
and competition coefficients.

\begin{table}
\centering\caption{Values of model parameters used for numerical integration.}\vspace{0.2cm}
\begin{tabular}{l|l|l}
Parameter     & Description & Value                                           \\ \hline
$R$         & innate growth rate, see \S\ref{ecoevo}  & $1.0$      \\
$\theta$    & abiotic optimum & $0.0$                              \\
$a$         & strength of abiotic selection & $0.01$               \\
$c$         & sensitivity to competition & $\{1.0\times10^{-7},5.0\times10^{-6}\}$ \\
$w$         & niche breadth & $0.1$                               \\
$U$         & total niche use & $1.0$                              \\
$\eta$      & segregation variance & $1.0$                         \\
$\mu$       & mutation rate & $1.0\times10^{-7}$                             \\
$V$ & variance of reproductive output & $5.0$
\end{tabular}\label{par_vals}
\end{table}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{community_dynamics}

}

\caption{\label{temporal}Temporal dynamics of mean trait (top), additive genetic variance (middle) and abundance (bottom) for the scenario of weak competition (left) and strong competition (right).}\label{fig:unnamed-chunk-6}
\end{figure}

\hypertarget{the-relation-between-the-strength-of-ecological-interactions-and-coevolution}{%
\subsection{\texorpdfstring{The relation between the strength of
ecological interactions and coevolution
\label{ecoevo}}{The relation between the strength of ecological interactions and coevolution }}\label{the-relation-between-the-strength-of-ecological-interactions-and-coevolution}}

Relating our treatment of resource competition to modern coexistence
theory (Chesson, 2000), the absolute competition coefficient
\(\alpha_{ij}\) becomes a dynamical quantity that can be written as

\begin{multline}
\alpha_{ij}(t)=\frac{c_i}{r_i(t)}\int_{-\infty}^{+\infty}\int_{-\infty}^{+\infty}p_i(x,t)p_j(y,t)\mathcal O_{ij}(x,y) dxdy \\ =\frac{c_iU_iU_j}{r_i(t)}\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right),
\end{multline}

where

\begin{equation}
r_i(t)=R_i-\frac{a_i}{2}\Big((\bar x_i(t)-\theta_i)^2+G_i(t)+\eta_i\Big).
\end{equation}

Hence, \(dN_i(t)\) can be expressed as

\begin{equation}
dN_i(t)=r_i(t)\left(1-\sum_{j=1}^S\alpha_{ij}(t)N_j(t)\right)N_i(t)dt+\sqrt{V_iN_i(t)}dW_1(t).
\end{equation}

Note that although \(r_i(t)\) is referred to in the coexistence
literature as the intrinsic growth rate of the population, \(R_i\) is a
deeper intrinsic quantity. For now we refer to \(R_i\) as the
\emph{innate} growth rate. Previous work has shown the importance of
demographic stochasticity (Schreiber, 2017; Turelli, 1980) and
evolutionary dynamics (Case and Taper, 1986; Schreiber et al., 2018) in
determining coexistence of competing species. However, these studies
consider evolution and demographic stochasticity separately. To our
knowledge, there are no rigorous theoretical investigations of
coexistence that account for evolutionary dynamics of quantitative
characters and demographic stochasticity simultaneously and, in
particular, no studies combining these processes that also include the
evolution of additive genetic variance. With the connection outlined in
this section formally established, researchers may pursue a postmodern
coexistence theory that naturally includes both the evolutionary
dynamics of quantitative characters and the effects of demographic
stochasticity in a simple synthetic framework.

In SM \S\ref{diffuse} we show that the standardized directional
selection gradient (sensu Lande and Arnold, 1983) induced by species
\(j\) on species \(i\) can be computed as

\begin{multline}
\beta_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big) \\
\times\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
\end{multline}

Our notation differs from Lande and Arnold (1983) in that subscripts
here denote species instead of components of multivariate traits and we
drop the prime that distinguishes between selection gradients and
standardized selection gradients.

\emph{Metric of pairwise coevolution:} Below we investigate the
correspondence of interaction intensity and coevolutionary change.
However, we can already identify one major discrepancy; \(\alpha_{ij}\)
is maximized when \(\bar x_i=\bar x_j\), but \(\beta_{ij}=0\) under the
same condition. We therefore include in our metric of selection the
standardized stabilizing selection gradient \(\gamma\) which measures
the effect of stabilizing or disruptive selection on phenotypic variance
(Lande and Arnold, 1983). In SM \S\ref{diffuse} we show that the
standardized stabilizing selection gradient induced by species \(j\) on
species \(i\) can be computed as

\begin{multline}
\gamma_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\left(1-b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big)^2\right) \\
\times\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
\end{multline}

To measure evolutionary change in species \(i\) induced by species
\(j\), we form the metric \(\Psi_{ij}=|\beta_{ij}|+|\gamma_{ij}|\) and
define the strength of coevolution as
\(\mathfrak{C}_{ij}=\Psi_{ij}\Psi_{ji}\). Figure \ref{net} displays the
joint distribution of \(\mathfrak{C}_{ij}\) and the product of
competition coefficients \(\alpha_{ij}\alpha_{ji}\) for two simulated
communities, both with richness \(S=1000\). The solid contour represents
the case of relatively high sensitivity to competition
(\(c=5\times10^{-7}\)) and the dashed contour represents the case of
relatively weak sensitivity to competition (\(c=1\times10^{-8}\)). In
both cases we see that \(\mathfrak{C}_{ij}\) and
\(\alpha_{ij}\alpha_{ji}\) are, essentially, unrelated. However, this
depiction is only representative of a specific set of parameters. Next,
we provide analytical approximations of the covariance between selection
gradients and competition coefficients and a numerical estimate for the
relationship between pairwise coevolution and competition coefficients
for a range of parameters.

\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{on_pl}

}

\caption{\label{net}Bivariate distributions of competition coffecients (y-axis) and pairwise coevolution (x-axis) under the scenarios of weak competition (dashed line) and strong competition (solid line) after simulating for $t=1000.0$ units of time. Simulations were ran for $S=1000$ species. Parameters are the same as in Table 2, except we used $c=1\times 10^{-8}$ for weak competition and $c=5\times 10^{-7}$ for strong competition in order to account for increased species richness.}\label{fig:unnamed-chunk-8}
\end{figure}

\emph{Covariance of selection and competition as a function of
diversity:} We now make use of the expressions derived for competition
coeffecients and selection gradients to investigate their relationship.
As a first pass, let us assume all model parameters are equivalent
across species and that each species has the same abundance and trait
variance. Let us further assume that richness \(S\) is large and the
distribution of mean trait values is normal with mean \(\bar{\bar x}\),
variance \(V_{\bar X}\) and density \(f_{\bar X}\). Such assumptions are
typical when deriving analytical results in the field of theoretical
coevolutionary community ecology (Nuismer et al., 2012, 2018). If
\(\bar{\bar x}\) is near \(\theta\) and \(V_{\bar X}\) is much smaller
than \(|2\frac{R}{a}-G-\eta|\), then we may approximate \(r_i\) with

\begin{multline}
\bar r=\int_{-\infty}^{+\infty}\left(R-\frac{a}{2}\Big((\bar x-\theta)^2+G+\eta\Big)\right)f_{\bar X}(\bar x)d\bar x \\
=R-\frac{a}{2}\Big((\bar{\bar x}-\theta)^2+V_{\bar X}+G+\eta\Big).
\end{multline}

In SM \S\ref{coef_grad_moms} we use these assumptions to calculate the
first and second order moments describing the joint distribution of
competition coefficients and selection gradients across the community.
We find that the covariance between linear selection gradients and
competition coeffecients are zero due to the symmetry implied by our
assumptions. However, setting \(\alpha(\bar x_i,\bar x_j)=\alpha_{ij}\),
\(\beta(\bar x_i,\bar x_j)=\beta_{ij}\) and
\(\gamma(\bar x_i,\bar x_j)=\gamma_{ij}\), the covariances between the
magnitude of linear selection gradients and competition coefficients and
between stabilizing selection gradients and competition coefficients can
be written as

\begin{subequations}\label{cov_eqns}
\begin{multline}\label{cov_alpha_beta}
\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)=\frac{2c^2b^2U^4N}{\pi\bar r}\sqrt\frac{V_{\bar X}}{2\pi} \\
\times\left(\frac{1}{(1+8bV_{\bar X})^{3/4}}-\frac{1}{(1+4bV_{\bar X})^{3/4}(1+2bV_{\bar X})^{1/2}}\right),
\end{multline}
\begin{equation}\label{cov_alpha_gamma}
\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)=\frac{c^2b^2U^4N}{2\pi\bar r}(1-2bV_{\bar X})\left(\frac{1}{\sqrt{1+4bV_{\bar X}}}-\frac{1}{1+2bV_{\bar X}}\right).
\end{equation}
\end{subequations}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/cov}

}

\caption{\label{cov_fig}Curves representing the covariance between the magnitude of linear selection gradients and competition coefficients (left) and between stabilizing selection gradients and competition coefficients (right) as a function of $2bV_{\bar X}$ which is approximately equal to the ratio of variance in mean traits among species to the intraspecific trait variance. In both plots we set $c=1.0\times10^{-4}$, $b=0.1$, $\bar r=0.1$ and $N=1.0\times10^{10}$ and let $V_{\bar X}$ vary.}\label{fig:unnamed-chunk-10}
\end{figure}

We visualize these relationships in Figure \ref{cov_fig} for fixed
\(c,b,\bar r,U\) and \(N\). In SM \S\ref{coef_grad_moms} we verify the
qualitative nature of these relationships by repeating numerical
integrations of system (\ref{comm_dynamics}) across a set of randomly
drawn model parameters. We designed our numerical approach to inherit
the assumptions of symmetric model parameters such as mutation rates
\(\mu_1=\dots=\mu_S\) and abiotic optima \(\theta_1=\dots=\theta_S\) and
to relax the key assumptions of normally distributed mean traits along
with constant additive genetic variances and abundances across species.

Our assumptions in this section imply \(b^{-1}=2(\sigma^2+w)\). If we
further assume \(\sigma^2+w\approx\sigma^2\), then
\(2bV_{\bar X}\approx V_{\bar X}/\sigma^2\). That is, when populations
are generalists and are comprised of specialist individuals, the value
\(2bV_{\bar X}\) is approximately equal to the ratio of interspecific
mean trait variation to intraspecific individual trait variation.
Glancing at Figure (\ref{cov_fig}) we see there is no relationship
between selection gradients and competition coefficients when this ratio
is zero. From equation (\ref{cov_alpha_beta}) we can use numerical
optimization to find that when \(V_{\bar X}/\sigma^2\approx1.25\) the
relationship between the magnitudes of linear selection gradients and
competition coefficients disappears, but when
\(V_{\bar X}/\sigma^2\lessapprox1.25\) (\(\gtrapprox1.25\)), this
covariance becomes negative (positive). Equation (\ref{cov_alpha_gamma})
states that when \(V_{\bar X}/\sigma^2\) is approximately equal to one
(or slightly larger), there is no expected relationship between
competition coefficients and quadratic selection gradients. However,
when \(V_{\bar X}/\sigma^2<1.0\) (\(>1.0\)), then we expect a positive
(negative) relationship between \(\alpha\) and \(\gamma\). These results
are true regardless of the chosen parameter values. In SM
\S\ref{coef_grad_moms} we use simulations of system
(\ref{comm_dynamics}) to show that these results do not qualitatively
differ when allowing for heterogeneous population sizes and additive
genetic variances across species.

From a biological perspective, if the ratio \(V_{\bar X}/\sigma^2\) is
small, then species are packed tightly in phenotypic space. In our model
this occurs when fitness is more sensitive to abiotic stabilizing
selection than competition (\(a\gg c\)). This causes species to overlap
more in niche space (i.e., large \(\alpha\)) and creates disruptive
selection for greater intraspecific variance (i.e., positive
\(\gamma\)), which explains the positive region of covariance between
\(\alpha\) and \(\gamma\). However, as species begin to overlap in niche
space, directional selection begins to vanish (i.e., small \(|\beta|\)),
leading to a negative covariance between \(\alpha\) and \(|\beta|\). In
the limiting case that two species have perfectly overlapping niches,
they will exhibit zero directional selection since a shift in either
direction will yield identical fitness advantages.

In the opposite scenario where fitness is much more sensitive to
competition than abiotic stabilizing selection (\(c\gg a\)), species
will not evolve to be as tightly packed and instead their niche-centers
will be spread out with little ovelap in their resouce utilization
curves (i.e., small \(\alpha\)). In this case biotic directional
selection will be strong (i.e., large \(|\beta|\)), particularly for
species towards the outer regions of niche space due to asymmetric
fitness advantages conferred by shifts in niche-centers. This leads to a
positive covariance between \(\alpha\) and \(|\beta|\). However, as
noted above, this directional selection will also erode away at standing
heritable variation (i.e., negative \(\gamma\)), reducing the rate at
which adaptation can occur and creating a negative covariance between
\(\alpha\) and \(\gamma\).

In summary, we see the relations between competition coefficients and
selection gradients are non-trivial and depend on the relative
magnitudes of different ecological processes shaping the community.
However, this does not address the relation between competition
coefficients and coevolution per se. In SM \S\ref{coef_grad_moms} we
show that calculating a formula for the covariance between competition
coefficients and the metric of coevolution \(\mathfrak{C}\) introduced
above provides a difficult analytical challenge. Instead of confronting
this challenge we build on our numerical approach used to justify
analytical approximations of
\(\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)\) and
\(\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)\) to approximate the
correlation of \(\alpha\) and \(\mathfrak{C}\). In Figure
\ref{cor_coev_fig} we plot values of
\(\mathrm{Corr}_{f_{\bar X}}(\alpha,\mathfrak{C})\) against the exact
ratio of interspecific trait variance to intraspecific trait variance.
Using a cubic regression, we see the correlation of coevolutionary
selection gradients and competition coefficeints is negative at variance
ratios below \(0.5\), zero at variance ratios between \(0.5\) and
\(1.0\), and drops below zero again at variance ratios above \(1.0\).

\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/Cαℭp}

}

\caption{\label{cor_coev_fig}Numerical estimates for the correlation of the strength of coevolution measured by $\mathfrak{C}$ and competition coefficeints $\alpha$ plotted against the variance of mean traits among species divided by the mean variance of traits within species. Each dot represents the result from a single simulation. The red line is a cubic regression.}\label{fig:unnamed-chunk-12}
\end{figure}

\hypertarget{conclusion}{%
\section{\texorpdfstring{Conclusion
\label{conclusion}}{Conclusion }}\label{conclusion}}

We have introduced an approach to derive models of evolutionary ecology
using the calculus of white noise and coined SAGA, a SPDE model of
phenotypic evolution that accounts for demographic stochasticity. From
SAGA we derived SDE that track the dynamics of abundance, mean trait and
additive genetic variance. Observing the expressions of these SDE we
find the effects of demographic stochasticity on the evolution of mean
trait and additive genetic variance characterize the effects of random
genetic drift. Although Lande (1976) has previously characterized the
effects of random genetic drift on mean trait evolution in quantitative
genetic models, the approach taken assumed constant effective population
size and discrete non-overlapping generations. In contrast, our approach
shows random genetic drift is a result of demographic stochasticity for
continuously reproducing populations with fluctuating abundances. To
illustrate the relevance of our approach to studies of evolutionary
ecology, we comined our SDE with classical competition theory to derive
a model of diffuse coevolution and used this model to investigate the
relationship between standardized selection gradients and competition
coefficients. We found the absolute value of linear selection gradients
are negatively related with competition coefficients when the ratio of
interspecific phenotypic variation to intraspecific phenotypic variation
is approximately less than \(1.25\) and positive when this ratio is
approximately greater than \(1.25\). We found the opposite relationship
between quadratic selection gradients and competition coefficients with
a threshold diversity ratio equal to unity instead of \(1.25\). This
work demonstrates that connecting contemporary theoretical approaches to
evolutionary ecology with some fundamental results in the theory of
measure-valued stochastic processes allows for the development of a
rigorous, yet flexible approach to synthesizing the dynamics of
abundance and distributions of quantitative characters, in agreement
with Champagnat, Ferrière and Méléard (2006). Although this approach
enjoys these important merits, there remains biological details and
their associated technical challenges that need to be confronted for
gaining a more thorough and accurate understanding of ecological
communities. We touch on just four of them here and provide suggestions
for future research to approach these challenges. For the first three,
we recommend the use of individual-based simulations of measure-valued
processes to confront these challenges. To summarize why, we state here
that analytical results become difficult to derive for small populations
and complicated models of mutation and selection.

\hypertarget{limitations-of-diffusion-limits-and-generalizing-coexistence-theory}{%
\subsection{Limitations of diffusion limits and generalizing coexistence
theory}\label{limitations-of-diffusion-limits-and-generalizing-coexistence-theory}}

As noted by Feller (1951), diffusion limits provide reasonable
approximations for large populations, but relatively small populations
require discrete models. Hence, as a diffusion limit, SPDE (\ref{SPDE})
restricts parameter values to regions that maintain large population
sizes. Although this suggests an important restriction on any model
developed under this framework by implying populations are not at risk
of extinction, the SDE describing abundance dynamics technically permits
extinction. However, for small abundances, pathology emerges in the
evolution of trait means and variances. In particular, stochastic
components of the SDE describing the evolution of \(\bar x\) and
\(\sigma^2\) diverge towards infinite values as \(N\to0\). Hence,
studies involving extinction and phenotypic evolution should be pursued
through a different approach. A natural alternative can be developed
utilizing the underlying measure-valued branching process that SAGA is a
diffusion limit of (see section \ref{stochastic}). This process
explicitly tracks individuals throughout their life-history and captures
reproduction as branching events. Hence, branching processes model
population size as a non-negative integer instead of a continuously
varying number. In particular, the pathological behavior described for
the SDE of \(\bar x\) and \(\sigma^2\) does not occur for measure-valued
branching processes. This concern relates to our comment in
\S\ref{ecoevo} on using our model of diffuse coevolution to extend
modern coexistence theory (sensu Chesson, 2000) in a way that accounts
for phenotypic evolution and demographic stochasticity. Just as previous
work has sought to understand the implications of processes for
coexistence separately, it seems necessary to determine whether or not
results pertaining to the coexistence of competing populations hold
using models that account for phenotypic evolution and demographic
stochasticity simultaneously. Because such investigations consider both
extinction and phenotypic evolution we argue the models used should be
either simulations of, or whenever possible, analytical results based on
measure-valued branching processes.

\hypertarget{genetic-architecture-and-intraspecific-trait-distributions}{%
\subsection{Genetic architecture and intraspecific trait
distributions}\label{genetic-architecture-and-intraspecific-trait-distributions}}

Our treatment of inheritance and our approach to model coevolution rests
on the assumptions of normally distributed breeding values and expressed
phenotypes along with asexual reproduction. However, real traits are not
encoded by an infinite number of loci each contributing an additive
infinitesimal effect (as required by the infinitesimal model), mutations
are not inherited as normally distributed deviations from parental
breeding values (as required by the Gaussian allelic model) and many
populations of interest exhibit non-random sexual reproduction.
Departures from this model of genetic architecture can produce more
complex distributions of breeding values and expressed traits. Such
deviations can be reinforced via strong non-Gaussian selection surfaces,
including the surface \(m(\nu,x)\) we have derived from niche theory,
along with non-random mating in sexually reproducing populations.
However, Gaussian approximations are convenient since they are defined
by their mean and variance. Future work investigating the effects of
non-normally distributed traits on the structure and dynamics of
ecological communities will therefore need to confront higher moments
such as skew and kurtosis, ideally integrating previously established
mathematical approaches to derive the dynamics of these higher moments
(Débarre et al., 2015).

Another approach to breaking the assumption of normally distributed
trait values is the development of explicit multilocus models. These
models describe the contributions of alleles at particular loci in the
genome to the development of quantitative traits. Tracking the
fluctuations of allele frequencies then allows theoreticians to
investigate the dynamics of phenotypic distributions that deviate from
normality. This approach has a long history in theoretical quantitative
genetics (Bulmer, 1980; Kirkpatrick et al., 2002; Turelli and Barton,
1994) and coevolutionary theory (Kopp and Gavrilets, 2006; Nuismer et
al., 2005, 2007). However, work to investigate the role of complex
genetic architectures in mediating feedbacks between the dynamics of
population abundances and the distributions of traits mediating
ecological interactions has apparently only just begun (Patel and
Bürger, 2019). Since analytical results are difficult to obtain for
general models of mutation, development and selection (c.f., Bürger,
2000), we again suggest the use of measure-valued branching processes
and their individual-based simulations. For example, by assuming the
mutational effects in the BBM model is Gaussian with variance an order
of magnitude greater than the equilibrium phenotypic (or additive
genetic) variance, we obtain the \emph{House-of-cards} model (Turelli,
1984). However, perhaps considering heavy-tailed distributions such as
the Laplace-distributions considered by Park et. al. (2017) will provide
more realistic yet less analytically tractable models of mutational
effects. Such mutation models have important implications in
conservation genetics where small populations evolve primarily in
response to mutation and random genetic drift (Lande, 1995).

\hypertarget{ecological-stoichiometry}{%
\subsection{Ecological stoichiometry}\label{ecological-stoichiometry}}

Our treatment of both biotic and abiotic selection neglects important
chemical constraints of biological reality. For instance, the resource
we assume species are competing over is modelled as a static quantity.
However, real resources can be dynamic quantities. Such theoretical
quantities may reflect abiotic cycles of material and energy or whole
trophic layers comprised of prey, hosts or mutualists. Although resource
dynamics have been captured theoretically in consumer-resource models
(Tilman, 1982), developing a more realistic model of resource
competition must incorporate details on the ecophysiology of organisms
(Loreau, 2010). Doing so will help clarify the feedback between the
evolution of populations and the ecosystem processes they are a part of.

Using plant-pollinator communities as an example, the role of nitrogen
mediating interspecific interactions has been reviewed by David, Storkey
and Stevens (2019) and the evolutionary ecology of the nutritional
content of nectar has been reviewed by Parachnowitsch, Manson and
Sletvold (2018). These studies demonstrate the need for further research
to understand how soil nutrient availability and organismal
ecophysiology structures communities of plants and pollinators.
Theoretical pursuits in this direction may profit from interfacing the
framework we have outlined here with population-ecosystem models such as
that developed by Fridley (2017). However, these population-ecosystem
models will likely yield complicated fitness functions that are
difficult to analyze using rigorous mathematical techniques. Hence, we
again recommend the use of simulations of measure-valued branching
processes that generalize BBM to understand when analytical
approximations hold.

\hypertarget{accounting-for-macroevolutionary-history}{%
\subsection{Accounting for macroevolutionary
history}\label{accounting-for-macroevolutionary-history}}

To understand patterns found in ecological communities considerations
must push beyond microevolutionary and contemporary ecological processes
and consider the macroevolutionary dynamics of ancestral lineages
leading to extant populations. Using sub-alpine flower communities as an
example, one can observe a very strict ordering of phenology across a
broad geographic range. In particular, whether in the Colorado Rocky
mountains (such as Gothic, Colorado) or on an outlier of the Idaho
batholith (such as Kamiak butte near Palouse, Washington), one will
almost surely observe a very conspicuous order of flowers emerging in
early spring: at the very beginning of the season blooms \emph{Claytonia
lanceolata} followed by \emph{Erythronium grandiflorum} which precedes
\emph{Delphinium nuttallianum} (B. Week, personal observations). If
contemporary phenological coevolution is rampant, why should this
pattern be so well preserved across a thousand miles of rugged and
diverse terrain? Although it would be exciting to find that these
species repeatedly coevolved this pattern in each location, a more
parsimoniuous hypothesis suggests the phenology and physiology of these
species slowly evolved independently over macroevolutionary time scales
to take advantage of the specific conditions available within each of
these windows of the flowering season. However, this could not have
carried out in the Rocky mountains since this terrain only became
habitable just over ten thousand years ago as the glaciers of the
Pleistocene began to retreat (Paul CaraDonna, personal communications).
Hence, given these considerations, it appears that an understanding of
early season phenology patterns should focus on how these communities
are assembled as opposed to contemporary evolutionary dynamics. Indeed,
recent work testing models of phylogeography ignores the potential for
contemporary evolution and instead suggests alpine flower communities
tend to follow neutral assembly where flowers merely compete for who can
disperse to new habitat first, as opposed to a selective process where a
regional species pool is filtered for those species adapted to the newly
available habitat (Marx et al., 2017).

Of course microevolutionary and ecological dynamics are not completely
irrelevant for understanding patterns in communities that are primarily
structured by deep evolutionary processes. In particular,
macroevolutionary trait evolution is simply the aggregation of
microevolutionary change occuring over large spans of time. This
suggests a road forward to connect the theory we have introduced to
models of macroevolutionary trait evolution.

Some approaches to modelling macroevolutionary trait change simply
repurpose microevolutionary models by blindly rescaling time from the
units of generations to millions of years (e.g., Nuismer and Harmon,
2014). Such an approach makes the implicit assumption that trait
evolution is statistically self-similar (sensu Falconer, 2014) so that
the stochastic evolution of traits on macroevolutionary time scales has
the same properties of trait evolution on microevolutionary time scales.
Although some stochastic processes, including Brownian motion, do
exhibit self-similarity, others do not. For example, consider a
modification of the Ornstein-Uhlenbeck process defined by the SDE

\begin{equation}
dX_t=a(\theta_t-X_t)dt+bdW_t
\end{equation}

where \(a,b>0\), \(W_t\) is a standard Brownian motion and \(\theta_t\)
is itself a Markov process that takes normally distributed jumps
centered on its current location at exponentially distributed time
intervals. If we assume the rate \(\lambda\) at which jumps occur is
much smaller than \(a\) and the variance in jumping is much larger than
\(b^2\), then, even though the sample paths of \(X_t\) are actually
continuous (if we zoom in close enough, they look like Brownian motion),
over long intervals of time sample paths of \(X_t\) will begin to appear
to have periods of continuity interrupted by an occasional discontinuous
jump and thus approach a qualitatively distinct process. These emergent
properties can be formally characterized by Lévy processes and have been
successfully employed in comparative phylogenetics to fit phenotypic
data from extant populations and the fossil record (Landis and
Schraiber, 2017). It would therefore be interesting to investigate
whether an application of a separation of time scales argument for the
rate of environmental change (\(\lambda\)) versus the rate of
evolutionary and ecological change (\(a\)) to microevolutionary models
derived via our framework can be used to obtain macroevolutionary models
that include not only mean trait evolution, but also the evolution of
trait variance and abundance. The resulting macroevolutionary models can
give rise to novel comparative phylogenetic methods and provide initial
conditions for microevolutionary models that capture contemporary
dynamics.

\hypertarget{final-remarks}{%
\section{Final remarks}\label{final-remarks}}

Although top-down approaches to community ecology, such as the Maximum
Entropy Theory of Ecology (Harte, 2011), have enjoyed some success in
describing community-level patterns (Harte and Newman, 2014; Xiao et
al., 2015), a mechanistic understanding of why these patterns emerge and
how they will change remains a formidable task. Such an understanding
must take both bottom-up and top-down approaches integrating
considerations from the ecophysiology of individual organisms that
reveal the economics of interspecific interactions (Sterner and Elser,
2008), to the phylogeographic history of taxa that sets the stage for
contemporary dynamics (Hickerson et al., 2010). Through connecting these
dots we can increase the variance explained in observations of
ecological communities by specific mechanisms and come closer to a
predictive theory of evolutionary community ecology. Despite the long
list of equations derived in this paper, this work takes just one small
step towards capturing these many details. However, we hope the
theoretical framework outlined here along with the demonstration of its
use in modelling competitive communities provides some helpful tools to
aid quantitative evolutionary ecologists in reaching such lofty goals.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-AbdalaRoberts2014}{}%
Abdala-Roberts, L., Mooney, K.A., 2014. Ecological and evolutionary
consequences of plant genotype diversity in a tri-trophic system.
Ecology 95, 2879--2893.

\leavevmode\hypertarget{ref-Athreya1972}{}%
Athreya, K.B., Ney, P.E., 1972. Branching processes. Springer Berlin
Heidelberg.

\leavevmode\hypertarget{ref-9781107648258}{}%
Ågren, G.I., Andersson, F.O., 2012. Terrestrial ecosystem ecology:
Principles and applications. Cambridge University Press.

\leavevmode\hypertarget{ref-Barton2019}{}%
Barton, N., Etheridge, A., 2019. Mathematical models in population
genetics.

\leavevmode\hypertarget{ref-Barton2013}{}%
Barton, N., Etheridge, A., Véber, A., 2013. Modelling evolution in a
spatial continuum. Journal of Statistical Mechanics: Theory and
Experiment 2013, P01002.

\leavevmode\hypertarget{ref-Barton2017}{}%
Barton, N., Etheridge, A., Véber, A., 2017. The infinitesimal model:
Definition, derivation, and implications. Theoretical Population Biology
118, 50--73.

\leavevmode\hypertarget{ref-Bertoin2003}{}%
Bertoin, J., Le Gall, J.-F., 2003. Stochastic flows associated to
coalescent processes. Probability Theory and Related Fields 126,
261--288.

\leavevmode\hypertarget{ref-Brown1956}{}%
Brown, W.L., Wilson, E.O., 1956. Character displacement. Systematic
Zoology 5, 49.

\leavevmode\hypertarget{ref-bulmer1980}{}%
Bulmer, 1980. The mathematical theory of quantitative genetics. Oxford
University Press.

\leavevmode\hypertarget{ref-Brger1986}{}%
Bürger, R., 1986. On the maintenance of genetic variation: Global
analysis of kimuras continuum-of-alleles model. Journal of Mathematical
Biology 24, 341--351.

\leavevmode\hypertarget{ref-9780471986539}{}%
Bürger, R., 2000. The mathematical theory of selection, recombination,
and mutation. Wiley.

\leavevmode\hypertarget{ref-Cantrell2004}{}%
Cantrell, R.S., Cosner, C., 2004. Spatial ecology via reaction-diffusion
equations. Wiley.

\leavevmode\hypertarget{ref-Case1986}{}%
Case, T.J., Taper, M.L., 1986. On the coexistence and coevolution of
asexual and sexual competitors. Evolution 40, 366--387.

\leavevmode\hypertarget{ref-Champagnat2006}{}%
Champagnat, N., Ferrière, R., Méléard, S., 2006. Unifying evolutionary
dynamics: From individual stochastic processes to macroscopic models.
Theoretical Population Biology 69, 297--321.

\leavevmode\hypertarget{ref-Chesson2000}{}%
Chesson, P., 2000. Mechanisms of maintenance of species diversity.
Annual Review of Ecology and Systematics 31, 343--366.

\leavevmode\hypertarget{ref-jeffreyconner2004}{}%
Conner, J.K., 2004. A primer of ecological genetics. Sinauer Associates.

\leavevmode\hypertarget{ref-9781932846126}{}%
Crow, J.F., Kimura, M., 1970. An introduction to population genetics
theory. The Blackburn Press.

\leavevmode\hypertarget{ref-Crutsinger2015}{}%
Crutsinger, G.M., 2015. A community genetics perspective: Opportunities
for the coming decade. New Phytologist 210, 65--70.

\leavevmode\hypertarget{ref-DaPrato2014}{}%
Da Prato, G., Zabczyk, J., 2014. Stochastic equations in infinite
dimensions. Cambridge University Press.

\leavevmode\hypertarget{ref-David2019}{}%
David, T.I., Storkey, J., Stevens, C.J., 2019. Understanding how
changing soil nitrogen affects plantpollinator interactions.
Arthropod-Plant Interactions 13, 671--684.

\leavevmode\hypertarget{ref-Dawson1975}{}%
Dawson, D.A., 1975. Stochastic evolution equations and related measure
processes. Journal of Multivariate Analysis 5, 1--52.

\leavevmode\hypertarget{ref-Dawson1978}{}%
Dawson, D.A., 1978. Geostochastic calculus. Canadian Journal of
Statistics 6, 143--168.

\leavevmode\hypertarget{ref-dawson1993measure}{}%
Dawson, D.A., 1993. Measure-valued markov processes, in: École d'été de
Probabilités de Saint-Flour Xxi-1991. Springer, pp. 1--260.

\leavevmode\hypertarget{ref-Dbarre2015EvolutionOQ}{}%
Débarre, F., Yeaman, S., Guillaume, F., 2015. Evolution of quantitative
traits under a migration-selection balance: When does skew matter? The
American naturalist 186 Suppl 1, S37--47.

\leavevmode\hypertarget{ref-Doebeli1996}{}%
Doebeli, M., 1996. Quantitative genetics and population dynamics.
Evolution 50, 532--546.

\leavevmode\hypertarget{ref-alisonetheridge2000}{}%
Etheridge, A.M., 2000. An introduction to superprocesses. American
Mathematical Society.

\leavevmode\hypertarget{ref-Etheridge2008}{}%
Etheridge, A.M., 2008. Drift, draft and structure: Some mathematical
models of evolution, in: Stochastic Models in Biological Sciences.
Institute of Mathematics Polish Academy of Sciences.

\leavevmode\hypertarget{ref-Etheridge1991}{}%
Etheridge, A., March, P., 1991. A note on superprocesses. Probability
Theory and Related Fields 89, 141--147.

\leavevmode\hypertarget{ref-lawrenceevans2010}{}%
Evans, L.C., 2010. Partial differential equations: Second edition.
American Mathematical Society.

\leavevmode\hypertarget{ref-9781470410544}{}%
Evans, L.C., 2014. An introduction to stochastic differential equations.
American Mathematical Society.

\leavevmode\hypertarget{ref-Evans1994}{}%
Evans, S.N., Perkins, E.A., 1994. Measure-valued branching diffusions
with singular interactions. Canadian Journal of Mathematics 46,
120--168.

\leavevmode\hypertarget{ref-Ewens2004}{}%
Ewens, W.J., 2004. Mathematical population genetics. Springer New York.

\leavevmode\hypertarget{ref-kennethfalconer2014}{}%
Falconer, K., 2014. Fractal geometry: Mathematical foundations and
applications. Wiley.

\leavevmode\hypertarget{ref-feller1951}{}%
Feller, W., 1951. Diffusion processes in genetics, in: Proceedings of
the Second Berkeley Symposium on Mathematical Statistics and
Probability. University of California Press, pp. 227--246.

\leavevmode\hypertarget{ref-Felsenstein1975}{}%
Felsenstein, J., 1975. A pain in the torus: Some difficulties with
models of isolation by distance. The American Naturalist 109, 359--368.

\leavevmode\hypertarget{ref-Fisher1923}{}%
Fisher, R.A., 1923. XXI.-on the dominance ratio. Proceedings of the
Royal Society of Edinburgh 42, 321--341.

\leavevmode\hypertarget{ref-Fitzpatrick2015}{}%
Fitzpatrick, C.R., Agrawal, A.A., Basiliko, N., Hastings, A.P., Isaac,
M.E., Preston, M., Johnson, M.T.J., 2015. The importance of plant
genotype and contemporary evolution for terrestrial ecosystem processes.
Ecology 96, 2632--2642.

\leavevmode\hypertarget{ref-Fitzpatrick2017}{}%
Fitzpatrick, C.R., Mikhailitchenko, A.V., Anstett, D.N., Johnson,
M.T.J., 2017. The influence of range-wide plant genetic variation on
soil invertebrate communities. Ecography 41, 1135--1146.

\leavevmode\hypertarget{ref-FRANK2012}{}%
Frank, S.A., 2012. Natural selection. IV. The price equation. Journal of
Evolutionary Biology 25, 1002--1019.

\leavevmode\hypertarget{ref-Fridley2017}{}%
Fridley, J.D., 2017. Plant energetics and the synthesis of population
and ecosystem ecology. Journal of Ecology 105, 95--110.

\leavevmode\hypertarget{ref-FUSSMANN2007}{}%
Fussman, G.F., Loreau, M., Abrams, P.A., 2007. Eco-evolutionary dynamics
of communities and ecosystems. Functional Ecology 21, 465--477.

\leavevmode\hypertarget{ref-Gomulkiewicz2017}{}%
Gomulkiewicz, R., Krone, S.M., Remien, C.H., 2017. Evolution and the
duration of a doomed population. Evolutionary Applications 10, 471--484.

\leavevmode\hypertarget{ref-Harmon2019}{}%
Harmon, L.J., Andreazzi, C.S., Débarre, F., Drury, J., Goldberg, E.E.,
Martins, A.B., Melián, C.J., Narwani, A., Nuismer, S.L., Pennell, M.W.,
Rudman, S.M., Seehausen, O., Silvestro, D., Weber, M., Matthews, B.,
2019. Detecting the macroevolutionary signal of species interactions.
Journal of Evolutionary Biology 32, 769--782.

\leavevmode\hypertarget{ref-Harte2011}{}%
Harte, J., 2011. Maximum entropy and ecology. Oxford University Press.

\leavevmode\hypertarget{ref-Harte2014}{}%
Harte, J., Newman, E.A., 2014. Maximum information entropy: A foundation
for ecological theory. Trends in Ecology \& Evolution 29, 384--389.

\leavevmode\hypertarget{ref-Hickerson2010}{}%
Hickerson, M., Carstens, B., Cavender-Bares, J., Crandall, K., Graham,
C., Johnson, J., Rissler, L., Victoriano, P., Yoder, A., 2010.
Phylogeography's past, present, and future: 10 years after avise, 2000.
Molecular Phylogenetics and Evolution 54, 291--301.

\leavevmode\hypertarget{ref-Hofbauer1998}{}%
Hofbauer, J., Sigmund, K., 1998. Evolutionary games and population
dynamics. Cambridge University Press.

\leavevmode\hypertarget{ref-Holt1987}{}%
Holt, R.D., 1987. On the relation between niche overlap and competition:
The effect of incommensurable niche dimensions. Oikos 48, 110.

\leavevmode\hypertarget{ref-Johnson2005}{}%
Johnson, T., Barton, N., 2005. Theoretical models of selection and
mutation on quantitative traits. Philosophical Transactions of the Royal
Society B: Biological Sciences 360, 1411--1425.

\leavevmode\hypertarget{ref-Kendall1966}{}%
Kendall, D.G., 1966. Branching processes since 1873. Journal of the
London Mathematical Society s1-41, 385--406.

\leavevmode\hypertarget{ref-Kimmel2015}{}%
Kimmel, M., Axelrod, D.E., 2015. Branching processes in biology.
Springer New York.

\leavevmode\hypertarget{ref-Kimura1965}{}%
Kimura, M., 1965. A stochastic model concerning the maintenance of
genetic variability in quantitative characters. Proceedings of the
National Academy of Sciences 54, 731--736.

\leavevmode\hypertarget{ref-Kimura1978}{}%
Kimura, M., Crow, J.F., 1978. Effect of overall phenotypic selection on
genetic change at individual loci. Proceedings of the National Academy
of Sciences 75, 6168--6171.

\leavevmode\hypertarget{ref-Kirkpatrick1727}{}%
Kirkpatrick, M., Johnson, T., Barton, N., 2002. General models of
multilocus evolution. Genetics 161, 1727--1750.

\leavevmode\hypertarget{ref-kolmogorov1999elements}{}%
Kolmogorov, A., Fomin, S., 1999. Elements of the theory of functions and
functional analysis. Dover.

\leavevmode\hypertarget{ref-Konno1988}{}%
Konno, N., Shiga, T., 1988. Stochastic partial differential equations
for some measure-valued diffusions. Probability Theory and Related
Fields 79, 201--225.

\leavevmode\hypertarget{ref-Kopp2006}{}%
Kopp, M., Gavrilets, S., 2006. Multilocus gentics and the coevolution of
quantitative traits. Evolution 60, 1321--1336.

\leavevmode\hypertarget{ref-Klzsch2015}{}%
Kölzsch, A., Alzate, A., Bartumeus, F., Jager, M. de, Weerman, E.J.,
Hengeveld, G.M., Naguib, M., Nolet, B.A., Koppel, J. van de, 2015.
Experimental evidence for inherent lévy search behaviour in foraging
animals. Proceedings of the Royal Society B: Biological Sciences 282,
20150424.

\leavevmode\hypertarget{ref-Kraft2007}{}%
Kraft, N.J.B., Cornwell, W.K., Webb, C.O., Ackerly, D.D., 2007. Trait
evolution, community assembly, and the phylogenetic structure of
ecological communities. The American Naturalist 170, 271--283.

\leavevmode\hypertarget{ref-Krylov1981}{}%
Krylov, N.V., Rozovskii, B.L., 1981. Stochastic evolution equations.
Journal of Soviet Mathematics 16, 1233--1277.

\leavevmode\hypertarget{ref-Lande1975}{}%
Lande, R., 1975. The maintenance of genetic variability by mutation in a
polygenic character with linked loci. Genetical Research 26, 221--235.

\leavevmode\hypertarget{ref-Lande1976}{}%
Lande, R., 1976. Natural selection and random genetic drift in
phenotypic evolution. Evolution 30, 314--334.

\leavevmode\hypertarget{ref-pmid17248993}{}%
Lande, R., 1980. The Genetic Covariance between Characters Maintained by
Pleiotropic Mutations. Genetics 94, 203--215.

\leavevmode\hypertarget{ref-Lande1982}{}%
Lande, R., 1982. A quantitative genetic theory of life history
evolution. Ecology 63, 607--615.

\leavevmode\hypertarget{ref-Lande1995}{}%
Lande, R., 1995. Mutation and conservation. Conservation Biology 9,
782--791.

\leavevmode\hypertarget{ref-Lande1983}{}%
Lande, R., Arnold, S.J., 1983. The measurement of selection on
correlated characters. Evolution 37, 1210.

\leavevmode\hypertarget{ref-Lande2003}{}%
Lande, R., Engen, S., SÆther, B.-E., 2003. Demographic and environmental
stochasticity, in: Stochastic Population Dynamics in Ecology and
Conservation. Oxford University Press, pp. 1--24.

\leavevmode\hypertarget{ref-Lande2009}{}%
Lande, R., Engen, S., Sæther, B.-E., 2009. An evolutionary maximum
principle for density-dependent population dynamics in a fluctuating
environment. Philosophical Transactions of the Royal Society B:
Biological Sciences 364, 1511--1518.

\leavevmode\hypertarget{ref-Landis2017}{}%
Landis, M.J., Schraiber, J.G., 2017. Pulsed evolution shaped modern
vertebrate body sizes. Proceedings of the National Academy of Sciences
114, 13224--13229.

\leavevmode\hypertarget{ref-9780691080628}{}%
Levins, R., 1968. Evolution in changing environments: Some theoretical
explorations. (MPB-2) (monographs in population biology). Princeton
University Press.

\leavevmode\hypertarget{ref-zeng1998absolute}{}%
Li, Z.-H., 1998. Absolute continuity of measure branching processes with
interaction. Chinese Journal of Applied Probability and Statistics 14,
231--242.

\leavevmode\hypertarget{ref-Lion2018}{}%
Lion, S., 2018. Theoretical approaches in evolutionary ecology:
Environmental feedback as a unifying perspective. The American
Naturalist 191, 21--44.

\leavevmode\hypertarget{ref-michelloreau2010}{}%
Loreau, M., 2010. From populations to ecosystems: Theoretical
foundations for a new ecological synthesis. Princeton University Press.

\leavevmode\hypertarget{ref-michaellynch1998}{}%
Lynch, M., Walsh, B., 1998. Genetics and analysis of quantitative
traits. Sinauer Associates is an imprint of Oxford University Press.

\leavevmode\hypertarget{ref-Arthur1969}{}%
MacArthur, R.H., 1969. Species Packing, and what Competition Minimizes.
Proceedings of the National Academy of Sciences 64, 1369--1371.

\leavevmode\hypertarget{ref-MacArthur1970}{}%
MacArthur, R.H., 1970. Species packing and competitive equilibrium for
many species. Theoretical Population Biology 1, 1--11.

\leavevmode\hypertarget{ref-9780691023823}{}%
MacArthur, R.H., 1972. Geographical ecology. Princeton University Press.

\leavevmode\hypertarget{ref-Macarthur1967}{}%
MacArthur, R.H., Levins, R., 1967. The limiting similarity, convergence,
and divergence of coexisting species. The American Naturalist 101,
377--385.

\leavevmode\hypertarget{ref-Manceau2016}{}%
Manceau, M., Lambert, A., Morlon, H., 2016. A unifying comparative
phylogenetic framework including traits coevolving across interacting
lineages. Systematic Biology syw115.

\leavevmode\hypertarget{ref-Marx2017}{}%
Marx, H.E., Dentant, C., Renaud, J., Delunel, R., Tank, D.C., Lavergne,
S., 2017. Riders in the sky (islands): Using a mega-phylogenetic
approach to understand plant species distribution and coexistence at the
altitudinal limits of angiosperm plant life. Journal of Biogeography 44,
2618--2630.

\leavevmode\hypertarget{ref-markmcpeek2017}{}%
McPeek, M.A., 2017. Evolutionary community ecology. Princeton University
Press.

\leavevmode\hypertarget{ref-DeMeester2018}{}%
Meester, L.D., Brans, K.I., Govaert, L., Souffreau, C., Mukherjee, S.,
Vanvelk, H., Korzeniowski, K., Kilsdonk, L., Decaestecker, E., Stoks,
R., Urban, M.C., 2018. Analyzing eco-evolutionary dynamics - the
challenging complexity of the real world. Functional Ecology.

\leavevmode\hypertarget{ref-meleard1992interacting}{}%
Méléard, M., Roelly, S., 1992. Interacting branching measure processes.
Stochastic Partial Differential Equations and Applications (G. Da Prato
and L. Tubaro, eds.) 246--256.

\leavevmode\hypertarget{ref-Mlard1993}{}%
Méléard, M., Roelly, S., 1993. Interacting measure branching processes.
Some bounds for the support. Stochastics and Stochastic Reports 44,
103--121.

\leavevmode\hypertarget{ref-Mubayi2019}{}%
Mubayi, A., Kribs, C., Arunachalam, V., Castillo-Chavez, C., 2019.
Studying complexity and risk through stochastic population dynamics:
Persistence, resonance, and extinction in ecosystems, in: Handbook of
Statistics. Elsevier, pp. 157--193.

\leavevmode\hypertarget{ref-9780674023383}{}%
Nowak, M.A., 2006. Evolutionary dynamics: Exploring the equations of
life. Belknap Press.

\leavevmode\hypertarget{ref-Nuismer2005}{}%
Nuismer, S.L., Doebeli, M., Browning, D., 2005. The coevolutionary
dynamics of antagonistic interactions mediated by quantitative traits
with evolving variances. Evolution 59, 2073.

\leavevmode\hypertarget{ref-Nuismer2014}{}%
Nuismer, S.L., Harmon, L.J., 2014. Predicting rates of interspecific
interaction from phylogenetic trees. Ecology Letters 18, 17--27.

\leavevmode\hypertarget{ref-Nuismer2012}{}%
Nuismer, S.L., Jordano, P., Bascompte, J., 2012. Coevolution and the
architecture of mutualistic networks. Evolution 67, 338--354.

\leavevmode\hypertarget{ref-Nuismer2007}{}%
Nuismer, S.L., Ridenhour, B.J., Oswald, B.P., 2007. Antagonistic
coevolution mediated by phenotypic differenes between quantitative
traits. Evolution 61, 1823--1834.

\leavevmode\hypertarget{ref-Nuismer2018}{}%
Nuismer, S.L., Week, B., Aizen, M.A., 2018. Coevolution slows the
disassembly of mutualistic networks. The American Naturalist 192,
490--502.

\leavevmode\hypertarget{ref-VanNuland2019}{}%
Nuland, M.E.V., Ware, I.M., Bailey, J.K., Schweitzer, J.A., 2019.
Ecosystem feedbacks contribute to geographic variation in plant-soil
eco-evolutionary dynamics across a fertility gradient. Functional
Ecology 33, 95--106.

\leavevmode\hypertarget{ref-PAGE2002}{}%
Page, K.M., Nowak, M.A., 2002. Unifying evolutionary dynamics. Journal
of Theoretical Biology 219, 93--98.

\leavevmode\hypertarget{ref-Parachnowitsch2018}{}%
Parachnowitsch, A.L., Manson, J.S., Sletvold, N., 2018. Evolutionary
ecology of nectar. Annals of Botany 123, 247--261.

\leavevmode\hypertarget{ref-Park2017}{}%
Park, B., Rutter, M.T., Fenster, C.B., Symonds, V.V., Ungerer, M.C.,
Townsend, J.P., 2017. Distributions of mutational effects and the
estimation of directional selection in divergent lineages ofArabidopsis
thaliana. Genetics 206, 2105--2117.

\leavevmode\hypertarget{ref-Parsons2010}{}%
Parsons, T.L., Quince, C., Plotkin, J.B., 2010. Some consequences of
demographic stochasticity in population genetics. Genetics 185,
1345--1354.

\leavevmode\hypertarget{ref-Patel2019}{}%
Patel, S., Bürger, R., 2019. Eco-evolutionary feedbacks between prey
densities and linkage disequilibrium in the predator maintain diversity.
Evolution 73, 1533--1548.

\leavevmode\hypertarget{ref-Perkins1991}{}%
Perkins, E.A., 1991. Conditional dawson-watanabe processes and
fleming-viot processes, in: Seminar on Stochastic Processes, 1991.
Birkhäuser Boston, pp. 143--156.

\leavevmode\hypertarget{ref-Perkins1992}{}%
Perkins, E.A., 1992. Measure-valued branching diffusions with spatial
interactions. Probability Theory and Related Fields 94, 189--245.

\leavevmode\hypertarget{ref-edwinarendperkins1995}{}%
Perkins, E.A., 1995. On the martingale problem for interactive
measure-valued branching diffusions. Amer Mathematical Society.

\leavevmode\hypertarget{ref-Polechov2015}{}%
Polechová, J., Barton, N.H., 2015. Limits to adaptation along
environmental gradients. Proceedings of the National Academy of Sciences
112, 6401--6406.

\leavevmode\hypertarget{ref-PRICE1970}{}%
Price, G.R., 1970. Selection and covariance. Nature 227, 520--521.

\leavevmode\hypertarget{ref-Queller2017}{}%
Queller, D.C., 2017. Fundamental theorems of evolution. The American
Naturalist 189, 345--353.

\leavevmode\hypertarget{ref-Reimers1989}{}%
Reimers, M., 1989. One dimensional stochastic partial differential
equations and the branching measure diffusion. Probability Theory and
Related Fields 81, 319--340.

\leavevmode\hypertarget{ref-Robertson1966}{}%
Robertson, A., 1966. A mathematical model of the culling process in
dairy cattle. Animal Science 8, 95--108.

\leavevmode\hypertarget{ref-joanroughgarden1979}{}%
Roughgarden, J., 1979. Theory of population genetics and evolutionary
ecology: An introduction. Macmillan.

\leavevmode\hypertarget{ref-Rudman2017}{}%
Rudman, S.M., Barbour, M.A., Csilléry, K., Gienapp, P., Guillaume, F.,
Jr, N.G.H., Hendry, A.P., Lasky, J.R., Rafajlović, M., Räsänen, K.,
Schmidt, P.S., Seehausen, O., Therkildsen, N.O., Turcotte, M.M., Levine,
J.M., 2017. What genomic data can reveal about eco-evolutionary
dynamics. Nature Ecology \& Evolution 2, 9--15.

\leavevmode\hypertarget{ref-Schreiber2017}{}%
Schreiber, S.J., 2017. Coexistence in the face of uncertainty, in:
Recent Progress and Modern Challenges in Applied Mathematics, Modeling
and Computational Science. Springer New York, pp. 349--384.

\leavevmode\hypertarget{ref-Schreiber2018}{}%
Schreiber, S.J., Patel, S., terHorst, C., 2018. Evolution as a
coexistence mechanism: Does genetic architecture matter? The American
Naturalist 191, 407--420.

\leavevmode\hypertarget{ref-Schuster1983}{}%
Schuster, P., Sigmund, K., 1983. Replicator dynamics. Journal of
Theoretical Biology 100, 533--538.

\leavevmode\hypertarget{ref-Skovmand2018}{}%
Skovmand, L.H., Xu, C.C., Servedio, M.R., Nosil, P., Barrett, R.D.,
Hendry, A.P., 2018. Keystone genes. Trends in Ecology \& Evolution 33,
689--700.

\leavevmode\hypertarget{ref-Sterner2008}{}%
Sterner, R., Elser, J., 2008. Ecological stoichiometry: Overview, in:
Encyclopedia of Ecology. Elsevier, pp. 1101--1116.

\leavevmode\hypertarget{ref-Taylor1978}{}%
Taylor, P.D., Jonker, L.B., 1978. Evolutionary stable strategies and
game dynamics. Mathematical Biosciences 40, 145--156.

\leavevmode\hypertarget{ref-davidtilman1982}{}%
Tilman, D., 1982. Resource competition and community structure.
Princeton University Press.

\leavevmode\hypertarget{ref-Turelli1980}{}%
Turelli, M., 1980. Niche overlap and invasion of competitors in random
environments II. The effects of demographic stochasticity, in:
Biological Growth and Spread. Springer Berlin Heidelberg, pp. 119--129.

\leavevmode\hypertarget{ref-Turelli1984}{}%
Turelli, M., 1984. Heritable genetic variation via mutation-selection
balance: Lerchs zeta meets the abdominal bristle. Theoretical Population
Biology 25, 138--193.

\leavevmode\hypertarget{ref-Turelli1986}{}%
Turelli, M., 1986. Gaussian versus non-gaussian genetic analyses of
polygenic mutation-selection balance, in: Evolutionary Processes and
Theory. Academic Press, pp. 607--628.

\leavevmode\hypertarget{ref-Turelli2017}{}%
Turelli, M., 2017. Commentary: Fisher's infinitesimal model: A story for
the ages. Theoretical Population Biology 118, 46--49.

\leavevmode\hypertarget{ref-pmid7851785}{}%
Turelli, M., Barton, N., 1994. Genetic and statistical analyses of
strong selection on polygenic traits: what, me normal? Genetics 138,
913--941.

\leavevmode\hypertarget{ref-Volpert2014}{}%
Volpert, V., 2014. Elliptic partial differential equations: Volume 2:
Reaction-diffusion equations. Springer Basel.

\leavevmode\hypertarget{ref-Walsh}{}%
Walsh, J.B., 1986. An introduction to stochastic partial differential
equations, in: Lecture Notes in Mathematics. Springer Berlin Heidelberg,
pp. 265--439.

\leavevmode\hypertarget{ref-Watanabe1968}{}%
Watanabe, S., 1968. A limit theorem of branching processes and
continuous state branching processes. Journal of Mathematics of Kyoto
University 8, 141--167.

\leavevmode\hypertarget{ref-Wright97}{}%
Wright, S., 1931. Evolution in mendelian populations. Genetics 16,
97--159.

\leavevmode\hypertarget{ref-Xiao2015}{}%
Xiao, X., McGlinn, D.J., White, E.P., 2015. A strong test of the maximum
entropy theory of ecology. The American Naturalist 185, E70--E80.

\leavevmode\hypertarget{ref-zheng2004nonlinear}{}%
Zheng, S., 2004. Nonlinear evolution equations. Chapman \& Hall/CRC
Press, Boca Raton, Fla.

\end{document}